Georgia's Financial Contributions to 2016 Presidential Campaigns
by Michael Harkin
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, Packages}
# Import R packages for project
library(ggplot2)
library(knitr)
library(dplyr)
library(gridExtra)
library(ggthemes)
```

The dataset I'm going to explore in this project comes from the FEC (Federal
Elections Commission). It contains data about individual financial contributions
from Georgia citizens to the 2016 United States presidential campaigns. I live
in Atlanta, generally considered a blue region in a predominantly red state.
Georgia went to Donald Trump in the 2016 election. I am curious what Georgia's
campaign contributions looked like in the most recent election. I am also
wondering what Atlanta's contributions looked like compared to the rest of the
state.

First I'll load in the data and take a look at the accompanying documentation
for 2016 datasets from the [FEC website](ftp://ftp.fec.gov/FEC/Presidential_Map/2016/DATA_DICTIONARIES/CONTRIBUTOR_FORMAT.txt). 

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the data and replace empty string values with NA
# Comma manually added to the end of the first line in order to avoid an error
ga <- read.csv('P00000001-GA.csv', na.strings=c(""))
# Drop the boolean "X" column that comes about when reading in the CSV file
ga$X <- NULL
```

# Univariate Plots Section
Now that the data is loaded in, I'll take a look at the names of the 18
variables in the dataset alongside the aforementioned documentation.

```{r echo=FALSE, message=FALSE, warning=FALSE, Names}
# Look at names of columns in ga data.frame
names(ga)
```

I'll also look at a couple of quick summaries of each column in the data, as 
well as look at the data's structure.

```{r echo=FALSE, message=FALSE, warning=FALSE, Data_Summary}
# Show summaries of all 18 columns in the 'ga' data.frame
summary(ga)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Data_Structure}
# Look at the structure of the 'ga' data.frame
str(ga)
```

Looking over these summaries, here are a few things I immediately wonder about:

* Why do some campaign contributions come in after Election Day?
* It'd be interesting to look at information about the top contributors.
* How do Atlanta contributions compare to those from the rest of the state?
* There are 35,000 more contributions to primaries than to general elections.
* Does the frequency of contributions change over time?

For ease of analysis later on, I'm going to convert the `contb_receipt_dt`
column to R's built-in Date format.

```{r echo=FALSE, message=FALSE, warning=FALSE, Convert_Rcpt_Dt_to_Date}
# Convert the contribution receipt date column to R's Date format
ga$contb_receipt_dt <- as.Date(ga$contb_receipt_dt, format='%d-%b-%y')
```

## Number of Contributions
As a starting point I'll take a look at a table containing the names of each 
candidate and the number of contributions they received, keeping in mind
that each of these totals include refunds.

```{r echo=FALSE, message=FALSE, warning=FALSE, Candidate_Table}
# Show table with names of each candidate in the 'ga' dataset
table(ga$cand_nm)
```

Unsurprisingly Trump and Hillary Clinton had the two highest counts of
contributions. I am surprised, however, that Bernie Sanders received only about
3,500 less contributions in Georgia than Trump, the eventual winner. This
conflicts with my prior assumptions about how red-leaning Georgia is.

## Amount of Contributions Received
In addition to the number of contributions associated with each candidate, it'd
be helpful to know the total amount of money received by each candidate.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contribution_Sums}
# Look at contribution dollar totals for each candidate by name
by(ga$contb_receipt_amt, ga$cand_nm, sum)
```

The highest totals belong to Clinton (7,045,095), Trump (4,935,612), Ted Cruz
(1,193,164), Sanders (1,075,672), and Ben Carson (842664.2). These totals
include both primary and general election contributions.

## Adding New Columns
I'd like to add the candidates' gender and party to the contribution records.
This additional information would potentially make for some interesting
analyses. In order to do this, I'll write out character vectors of candidates
within each political party based on information I collected from the [FEC](http://www.fec.gov/press/resources/2016presidential_form2nm.shtml).

I'll use nested ifelse statements to assign each contribution to the
appropriate political party. I'll also do the same kind of categorization using
character vectors to match the candidate's gender to each contribution.

Following the format of variables as they came titled, I'll name these new
columns `cand_party` (candidate party) and `cand_gender` (candidate gender).

```{r echo=FALSE, message=FALSE, warning=FALSE, Add_Gender_and_Party}
# Create character vectors of parties each candidate is in
democratic <- c("Clinton, Hillary Rodham", "Lessig, Lawrence",
                "O'Malley, Martin Joseph", "Sanders, Bernard", 
                "Webb, James Henry Jr.")
green <- c("Stein, Jill")
independent <- c("McMullin, Evan")
libertarian <- c("Johnson, Gary")
republican <- c("Bush, Jeb", "Carson, Benjamin S.", "Christie, Christopher J.",
                "Cruz, Rafael Edward 'Ted'", "Fiorina, Carly", 
                "Graham, Lindsey O.", "Huckabee, Mike", "Jindal, Bobby",
                "Kasich, John R.", "Paul, Rand", "Perry, James R. (Rick)",
                "Rubio, Marco", "Santorum, Richard J.", "Trump, Donald J.",
                "Walker, Scott")

# Create character vectors of male and female candidates
male_candidates <- c("Bush, Jeb", "Carson, Benjamin S.",
                     "Christie, Christopher J.", "Cruz, Rafael Edward 'Ted'",
                     "Graham, Lindsey O.", "Huckabee, Mike", "Jindal, Bobby",
                     "Johnson, Gary", "Kasich, John R.", "Lessig, Lawrence",
                     "McMullin, Evan", "O'Malley, Martin Joseph", "Paul, Rand",
                     "Perry, James R. (Rick)", "Rubio, Marco",
                     "Sanders, Bernard", "Santorum, Richard J.",
                     "Trump, Donald J.", "Walker, Scott",
                     "Webb, James Henry Jr.")
female_candidates <- c("Clinton, Hillary Rodham", "Fiorina, Carly",
                       "Stein, Jill")

# Nested ifelse statements to assign correct candidate party name to each row
ga$cand_party <- ifelse(ga$cand_nm %in% republican, "Republican", 
                    ifelse(ga$cand_nm %in% democratic, "Democrat",
                      ifelse(ga$cand_nm %in% green, "Green",
                        ifelse(ga$cand_nm %in% libertarian, "Libertarian",
                            ifelse(ga$cand_nm %in% independent, "Independent", 
                              NA)))))
# Nested ifelse statements to assign correct candidate gender to each row
ga$cand_gender <- ifelse(ga$cand_nm %in% male_candidates, "male",
                    ifelse(ga$cand_nm %in% female_candidates, "female", NA))
```

I'll take a quick look at summaries and plots of these new variables.

```{r echo=FALSE, message=FALSE, warning=FALSE, Gender_Table}
# Create table showing counts of contributions to candidates by gender
table(ga$cand_gender)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Gender_Bar_Chart}
# Create bar chart showing counts of contributions to candidates by gender
qplot(x = cand_gender, data = ga)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Party_Table}
# Create table showing counts of contributions to each political party
table(ga$cand_party)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Candidate_Parties_ScaleYLog10}
# Create bar chart showing number of contributions to each party
qplot(x = cand_party, data = ga) +
  # Apply a log10 transformation to the y-axis and remove scientific notation
  scale_y_log10(labels = scales::comma)
```

The Democratic candidates received over 35,000 more contributions than the
Republican candidates, despite there being 15 Republican candidates during the
primary elections. Additionally, there are only 15,000 more contributions
associated with male candidates than female candidates, even though there were
20 male candidates and 3 female candidates. 

The third-party candidates are barely perceptible in this bar graph, so I'll
apply a log10 scale to the y-axis in order to get a better sense of the 
distribution.

Looking at this data on a log10 y-scale makes it more apparent that the amount
of contributions to third-party candidates is not insignificant. Gary Johnson,
the Libertarian candidate, received 451 contributions and Jill Stein, the Green
candidate, received 157 contributions. The bar chart without the log10
transformation makes it look like the Independent candidate, Evan McMullin, has
0 contributions, whereas he actually has 41.

## Contributions by Candidate
I'd like to get a better picture of how contributions to the individual
candidates break down. First I'll take a look at a bar chart of the Democratic
candidates. I'll apply a log10 transformation to the y-axis in order to be able
to properly view contributions to all the candidates. I'll also swap the x-axis
and y-axis for ease of reading.

```{r echo=FALSE, message=FALSE, warning=FALSE, Democratic_Contributions_ScaleYLog10}
# Create bar chart showing number of contributions to Democratic candidates
qplot(x = cand_nm, data = subset(ga, cand_party == 'Democrat')) +
  # Apply log10 transformation to y-axis and remove scientific notation
  scale_y_log10(labels = scales::comma) +
  # Swap x- and y-axes for ease of reading
  coord_flip()
```

Clinton, the Democratic nominee, received considerably more donations in Georgia
than any other Democratic candidate, getting around 40,000 more contributions
than Sanders, the runner-up.

The other Democratic candidates received very few contributions. Here are the
contribution counts for the other three Democrats:

* Jim Webb: 14
* Lawrence Lessig: 16
* Martin O'Malley: 17

Now I'll turn to the Republicans. I'll do the same log10 transformation to the
y-axis in order to be able to see all the contributions properly, and will
swap the x- and y-axis for ease of reading.

```{r echo=FALSE, message=FALSE, warning=FALSE, Republican_Contributions_ScaleYLog10}
# Create bar chart showing all the Republican candidates
qplot(x = cand_nm, data = subset(ga, cand_party == 'Republican')) +
  # Apply a log10 transformation to the y-axis and remove scientific notation
  scale_y_log10(labels = scales::comma) +
  # Swap the x- and y-axis for ease of reading
  coord_flip()
```

There's significant variance in the amount of contributions received by the
Republican candidates. The four highest bars represent the number of
contributions received by Trump, Ted Cruz, Ben Carson, and Marco Rubio. 

As for the lower-ranking candidates, here are their contribution counts:

* Mike Huckabee: 291
* Scott Walker: 202
* Lindsey Graham: 75
* Chris Christie: 45
* Rick Perry: 27
* Rick Santorum: 17
* Bobby Jindal: 12

I'd also like to look at the candidates outside of the two major political
parties.

```{r echo=FALSE, message=FALSE, warning=FALSE, Independent_Contributions}
# Create bar chart showing counts of contributions to third-party candidates
qplot(x = cand_nm, data = subset(ga, cand_party != 'Democrat'
                                 & cand_party != 'Republican'))
```

These three candidates received much fewer contributions than the top Democratic
and Republican candidates did. Gary Johnson, the Libertarian candidate and the
recipient of the highest number of contributions among the third-party
candidates, has just over 450 contributions associated with his campaign.

## Contribution Amounts
I also need to take a look at contribution receipt amounts. First I will run a
quick summary of the `contb_receipt_amt` variable, which represents the receipt
amount for each contribution, in order to get a sense of the range of values of
contributions received.

```{r echo=FALSE, message=FALSE, warning=FALSE, Receipts_Summary}
# Show summary of contribution receipt amounts for entire dataset
summary(ga$contb_receipt_amt)
```

There are at least a few negative values in the dataset, and there appear to be
both positive and negative outliers. The negative values represent refunds to
contributors. According to the summary of the dataset's `receipt_desc` variable,
there are 1,047 rows of the dataset that are marked as "Refund." I'll keep this
in mind as I continue exploring the data.

According to [FEC rules](http://www.fec.gov/pages/brochures/citizens.shtml), an
individual maximum of 2,700 dollars can be contributed to a primary or general
presidential election. It's legal to contribute as much as 5,400 dollars to a
campaign if half the amount is designated for the primary election and the
other half is designated to the general election. For this reason I'm curious
about the significant outliers in this data, particularly the low of -5,900
dollars and the high of 12,500 dollars.

While looking at the overall distribution of contributions, it'd be helpful to
look at a histogram of the range of receipt amounts. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Receipts_Histogram}
# Create histogram showing counts of campaign contribution amounts
qplot(x = contb_receipt_amt, data = ga)
```

Most contributions to the various presidential campaigns appear to be less than
500 dollars per this initial `qplot` graph. There are some real outliers in the
data as well, particularly a handful of very low negative values and some very
high outliers as well.

I'm going to take a closer look at the tallest chunk of the distribution,
adjusting the binwidth and the axes to get a sense of the distribution of
contributions under 500 dollars.

```{r echo=FALSE, message=FALSE, warning=FALSE, Receipts_Hist_Zoom}
# Create histogram showing counts of contribution amounts between 0 and 500
qplot(x = contb_receipt_amt, data = ga, binwidth = 25, xlim = c(0, 500))
```

The data remains positively skewed even zoomed in this far. I can now see that
most contributions are 100 dollars or less. I'm going to apply a transformation
to the data using `scale_x_sqrt` to see if that'll give a helpful view of the
distribution's shape. I'll also reduce the binsize to 1.

```{r echo=FALSE, message=FALSE, warning=FALSE, Receipts_Hist_Sqrt_Transform}
# Plot a histogram showing counts of the contribution receipt amounts
# Add 1 to contribution receipt amount in order to include 0 values
qplot(x = (contb_receipt_amt + 1), data = ga) +
  scale_x_sqrt()
```

The square root transformation results in a simlarly positively skewed
distribution. I'll try a log10 transformation of the data to see if that's
helpful.

```{r echo=FALSE, message=FALSE, warning=FALSE, Receipts_Hist_Log_Transform}
# Apply a log10 scale to the contribution amounts
# Add 1 to contribution receipt amount in order to include zeroes
qplot(x = (contb_receipt_amt + 1), data = ga) +
  scale_x_log10()
```

A log10 transformation gives an approximately normal distribution of the data
for contribution amount. However, because it's a log transformation, it doesn't
incorporate the negative values. Same goes for the square root transformation.
That said, the refunds make up only a small portion of the overall data (less
than 1 percent).

I'd like to zoom in on the lower end of the distribution (minus the log10
transformation) to get a better idea of how much there were of certain
contribution amounts.

```{r echo=FALSE, message=FALSE, warning=FALSE, Receipts_Hist_250}
# Create histogram of contributions between 0 and 250 with binwidth of 5
qplot(x = contb_receipt_amt, data = ga, binwidth = 5, xlim = c(0, 250))
```

Many of these contributions are below 50 dollars, so I'd like to zoom in a bit
more on the part of the histogram between 0 and 50 dollars.

```{r echo=FALSE, message=FALSE, warning=FALSE, Receipts_Hist_50}
# Create histogram of contributions below 50 dollars with binwidth 1
qplot(x = contb_receipt_amt, data = ga, binwidth = 1, xlim = c(0, 50))
```

Most of the contributions are multiples of 5, with the highest number of
contributions coming in at 25 dollars. There are a high number of 5-, 10-, 50-,
and 100-dollar contributions as well.

## Contribution Dates
Another potentially interesting component of the data are the dates on which
contributions were received. Earlier I converted the `contb_receipt_dt` column
to the R date format, which should help with displaying and analyzing the data
correctly.

First I'll do a histogram of receipt dates for every contribution in Georgia.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contrib_Receipt_Date_Hist}
# Create histogram of contribution counts for each date in the dataset
qplot(x = contb_receipt_dt, data = ga)
```

There's a surprisingly significant amount of contributions in the year 2015, and
even a few in 2014. The contributions ramp up quite a bit in 2016. It makes
sense that the majority of contributions would come in during 2016 given that
presidential campaigns don't seem to ramp up until the year of the election
itself.

I'm curious what this distribution looks like with the binwidth set to 1 and the
x-axis confined to 2015 through 2016.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contrib_Receipt_Dt_Adjusted}
# Create histogram of contribution counts for each date in the dataset
# Set binwidth to 1
qplot(x = contb_receipt_dt, data = ga, binwidth = 1) +
  scale_x_date(lim = c(as.Date('2015-01-01'), as.Date('2016-12-31')))
```

This results in a much more granular histogram with distinct peaks around March
2016, July 2016, and just before Election Day in November 2016. In the
multivariate analysis I want to be sure to examine whether the number of
contributions and their amounts have any relationship with the primary and
general elections.

## Form Type
There are three form types associated with each contribution in the dataset. 
These categorizations refer to lines in Form 3P, an FEC document that requires
candidates to report receipts and disbursements. Here are the descriptions
of each line:

**Form 3P Line 17A**: Contributions (other than loans) from Individuals/Persons
Other than Political Committee

**Form 3P Line 18**: Transfers from Other Authorized Committees

**Form 3P Line 28A**: Refunds of Contributions to Individuals/Persons Other Than
Political Committees

Here's a quick look at the amount of each form type that appears in the dataset
for Georgia.

```{r echo=FALSE, message=FALSE, warning=FALSE, Form_Type_Table}
# Display table showing counts of contribution form types
table(ga$form_tp)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Form_Type_Bar_Chart}
# Create bar chart showing counts of contribution form types
qplot(x = form_tp, data = ga)
```

The vast majority of contributions (116,284) are categorized as "SA17A", meaning
they are contributions from individuals. There are a further 36,654 that are
contributions from committees. It'd be interesting in the bivariate section to
see if these form types have any type of correspondence with donation amounts or
contributor location.

## Contributor Names
I'm curious about the names listed as having hundreds of contributions
associated with their name. The table listing out every contributor's name is
far too long to print out but below are a few contributors who have many
transactions associated with their name.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contributor_Summary}
# Show the contributors with the highest number of contributions associated
# with their name.
head(summary(ga$contbr_nm))
```

I'll look at the sum of the top three of these folks' contributions.

Julie Ralston:

```{r echo=FALSE, message=FALSE, warning = FALSE, Indiv_Contbr_Summary_1}
# Create subset of data for contributions by Julie Ralston
julie_ralston <- subset(ga, contbr_nm == "RALSTON, JULIE")
# Display sum of Julie Ralston's campaign contributions
sum(julie_ralston$contb_receipt_amt)
```

Joe Tate:

```{r echo=FALSE, message=FALSE, warning = FALSE, Indiv_Contbr_Summary_2}
# Create subset of data for contributions by Joe Tate
joe_tate <- subset(ga, contbr_nm == "TATE, JOE")
# Display sum of Joe Tate's campaign contributions
sum(joe_tate$contb_receipt_amt)
```

Janice Gladin:

```{r echo=FALSE, message=FALSE, warning = FALSE, Indiv_Contbr_Summary_3}
# Create subset of data for contributions by Janice Gladin
janice_gladin <- subset(ga, contbr_nm == "GLADIN, JANICE")
# Display sum of Janice Gladin's campaign contributions
sum(janice_gladin$contb_receipt_amt)
```

None of these three contributors pass the 2,700-dollar mark. I'm wondering if
there was perhaps a program that allowed people to contribute a certain amount
of money per week or per month on a recurring basis. That might help explain
the recurring payments in cases like this.

In the bivariate plots section, I want to look at individual contributor totals
if possible.

## Memo Text
One of the dataset's columns is named `memo_cd`, and it indicates whether or not
there is "memo text"" associated with the contribution. If an 'X' appears in
this column then there is an explanatory memo of some kind associated with that
contribution.

```{r echo=FALSE, message=FALSE, warning=FALSE, Memo_Code_Table}
# Display a table of the values found in the 
table(ga$memo_cd)
```

Almost 38,000 of the contributions have "memo text" associated with them, and
there are over 100 distinct memos. Here's a sample of the memos that appear.

```{r echo=FALSE, message=FALSE, warning=FALSE, Memo_Text_Table}
# Show the first few rows of a table containing every possible memo in the data
head(table(ga$memo_text))
```

Here are a few of the notable categories from the extended list:

* 25,549 labeled as "Earmarked Contribution"
* 12,383 labeled as "Hillary Victory Fund"
* 227 labeld as "Earmarked from Make DC Listen"
* 167 labeled as "Best Efforts Update"
* Several labeled as a refund or redesignation of some kind

## Election Type
The `election_tp` column specifies which election each contribution is 
connected to. The vast majority of contributions will likely be associated with
the 2016 election but I'd like to see what the data says.

```{r echo=FALSE, message=FALSE, warning=FALSE, Election_Type_Table}
# Show a table of the counts of the different election types in the dataset
table(ga$election_tp)
```

Nearly 95,000 the contributions are associated with primaries (P2016) and a
further 59,000 are associated with the general election (G2016). There are 23
labeled as "other" (O2016) - it'd be interesting to take a look at those - and
one contribution labeled as for the 2020 primaries (P2020).

I'll look at counts of election type and use a log10 scale on the y-axis in
order to properly see the counts.

```{r echo=FALSE, message=FALSE, warning=FALSE, Election_Type_Bar_Chart_YLog10}
# Create bar chart showing different election types in the dataset
qplot(x = election_tp, data = ga) +
  # Apply a log10 transform to the y-scale and remove scientific notation
  scale_y_log10(label = scales::comma)
```

There are a significant amount of 'NA' values for election type. I'll look at
these contributions shortly.

First I'll look at the contributions marked as "O2016" (2016 "Other") and the
contribution marked as "P2020" (for the 2020 primaries).

2016 "Other":

```{r echo=FALSE, message=FALSE, warning=FALSE, Other_2016}
# Create subset of data where election type is labeled as "Other"
other_2016 <- subset(ga, election_tp == "O2016")
head(other_2016)
```

2020 Primaries:

```{r echo=FALSE, message=FALSE, warning=FALSE, Primary_2020}
# Create subset of data where election type is labeled as "2020 Primary"
primary_2020 <- subset(ga, election_tp == "P2020")
head(primary_2020)
```

The contributions labeled "other" are all donations to Jill Stein, the Green
nominee, made between November 23-28, 2016.

The donation labeled as for the 2020 primaries was made to Lindsey Graham on
January 27, 2016 and is labeled as "Redesignation from General."

Now I'd like to take a look at the rows where the election type is "NA."

```{r echo=FALSE, message=FALSE, warning=FALSE, NA_2016}
# Create subset of data where election type is labeled as "NA"
na_2016 <- subset(ga, is.na(election_tp))
```

Number of rows with election type "NA":

```{r echo=FALSE, message=FALSE, warning=FALSE, NA_2016_Dimensions}
# Count the number of rows with election type "NA"
dim(na_2016)
```

It turns out that 459 of the transactions have their election type labeled as
"NA." I'm going to take a closer look at these contributions.

```{r echo=FALSE, message=FALSE, warning=FALSE, NA_2016_Summary}
summary(na_2016)
```

The vast majority of the transactions with election type "NA" - 408 of them -
are to Trump's campaign. 408 of them have the form type "SA18" meaning they are
contributions from authorized committees. Additionally, about one-third of the
contributions are from people who are retired, and most of them are from parts
of Georgia outside Atlanta.

I'll filter the data.frame using `dplyr` to look more closely at contributions
with election type "NA" and form type "SA18".

```{r echo=FALSE, message=FALSE, warning=FALSE, NA_and_SA18_Dataframe_Summary}
# Filter ga.frame to election type "NA" and form type "SA18"
na_sa18 <- filter(ga, is.na(election_tp), form_tp == "SA18")
# Display summary of this new data.frame
summary(na_sa18)
```

All of the "NA" contributions labeled as transfers from authorized committees
are contributions to Donald Trump that came in between August 16 and November
17, 2016. Almost half of these contributors are either retired, self-employed,
or homemakers, and most of them live outside of Atlanta.

# Univariate Analysis

### What is the structure of your dataset?
There are 153,985 observations of 18 variables. Each row in the dataset
represents a contribution, refund, or redesignation from Georgia individuals to
the 2016 presidential campaigns. This data is drawn from reports from each
campaign as required by the Federal Election Commission (FEC).

The individual transactions have several columns' worth of information. This
information includes the amount and date of each transaction as well as the
candidate it was made in support of, whether the transaction was for the primary
or general election, and demographic information about the contributor. In cases
where an individual had multiple transactions associated with their name, each
individual transaction is recorded, as opposed to being consolidated into one
row per contributor.

### What is/are the main feature(s) of interest in your dataset?
The main features of interest to me are the contribution amounts, the dates and
designations of each contribution, and the demographic information about the
contributors. The way this data is reported allows for several potential avenues
of exploration. It gets a bit fuzzy with regard to the wide range of memos and
designations that are used.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
I think that the zip code, form type, and individual transaction IDs will help
with reshaping and mutating the data in order to better understand patterns and
trends in the contribution data. These parts of the data should be helpful with
analyzing contributions by location.

### Did you create any new variables from existing variables in the dataset?
I added two new variables: candidate party (`cand_party`) and gender
(`cand_gender`). I used character vectors of the candidates' names to do this.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

There are some significant outliers in the distribution of contribution amounts.
These outliers positively skew the contribution distributions. For instance,
there's a Republican contribution that's 12,500 dollars, as well as one refund
to a Republican contributor in the amount of -5,900 dollars.

I transformed the contribution amount data using a log10 scale. I did this
because I wanted to determine whether the distribution was approximately
normal, and it appeared to be positively skewed prior to transformation. The
log10 transformation resulted in normal-ish distributions for all of the
candidates I looked at who received at least a couple hundred contributions.

There were a few bar charts with which I applied a log10 transformation to the
y-axis. This allowed me to get a better sense of the distribution in cases where
the counts for individual variables were hard to perceive or seemed to be zero.

I also changed the contribution date column to R's Date format in order to be
able to work with that field in a more convenient way that accurately reflects
that component of the data set. 

# Bivariate Plots Section

## Contribution Amounts by Party
I'd be interested in seeing what the individual parties' contributions look
like, particularly the Republicans and Democrats given that there is a decent
sample size of contributions for each party.

I'll start by using the `by` function to get a summary of contribution amounts
divided up by party.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Amts_by_Party}
# Summarize receipt amounts by candidates' political parties
by(ga$contb_receipt_amt, ga$cand_party, summary)
```

The Democrats and Republicans have the lowest median contribution amounts at 25
dollars and 50 dollars respectively. Republicans, however, have a much higher
mean contribution of 158 dollars compared to a mean contribution of 86 dollars
to the Democratic candidates. It seems notable that the Republican candidates'
mean is that much higher given the high number of Republican candidates who ran
and are represented in the dataset.

Additionally, the median and mean contribution amounts for candidates outside of
the two major political parties were considerably higher than the mean and
median contributions to Democratic candidates. This is interesting for a couple
of reasons. Perhaps some contributors to third-party candidates and nominees
pledge more money because they're aware that there are fewer third-party
campaign contributors than there are contributors to Democratic and Republican
candidates and nominees.

A frequency polygon might be a useful way of looking at the distribution of
contribution amounts across the different U.S. political parties. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Contbs_by_Party_FreqPoly}
# Create frequency polygon showing contribution amounts by party
qplot(x = contb_receipt_amt, data = ga, binwidth = 1, geom = 'freqpoly',
      colour = cand_party)
```

Every perceptible peak in this frequency polygon represents either Democratic or
Republican candidates. Let's zoom in to the area between 0 and 250 dollars, and
then further in at the area between 0 and 100 dollars, just to get a sense of
the distribution of the lower contribution amounts that comprise the bulk of
contributions received in this election. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Contbs_Party_FreqPoly_250}
# Create frequency polygon showing contribution amounts under 250 by party
qplot(x = contb_receipt_amt, data = ga, binwidth = 1, geom = 'freqpoly',
      colour = cand_party) +
  scale_x_continuous(limits = c(0, 250))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, FreqPoly_Parties_100}
# Create frequency polygon showing contribution amounts under 100 by party
qplot(x = contb_receipt_amt, data = ga, binwidth = 1, geom = 'freqpoly',
      colour = cand_party) +
  scale_x_continuous(limits = c(0, 100), breaks = seq(0, 100, 5))
```

It looks like the Democratic candidates had generally much higher levels of
giving at amounts of 100 dollars or less. There's a spike at the 40-dollar and
80-dollar levels for Republican candidates, as well as spikes at the 200- and
250-dollar levels. However, nearly every other spike for Republican candidates
is dwarfed by a larger Democratic candidates spike above it. The Independent,
Green, and Libertarian candidates barely register on this graph, appearing to
be a single flat line just above a count of 0.

## Contribution Amounts by Candidate
Now I'll take a look at histograms of contributions for the various candidates.
First I'll look at the top two Democratic candidates.

Hillary Clinton:

```{r echo=FALSE, message=FALSE, warning=FALSE, Clinton_Contributions_Summary}
# Create subset of data for contributions to Clinton
clinton <- subset(ga, cand_nm == 'Clinton, Hillary Rodham')
# Display summary of this new subset
summary(clinton$contb_receipt_amt)
```

Bernie Sanders:

```{r echo=FALSE, message=FALSE, warning=FALSE, Sanders_Contribution_Summary}
# Create subset of data for contributions to Sanders
sanders <- subset(ga, cand_nm == 'Sanders, Bernard')
# Display summary of this new subset
summary(sanders$contb_receipt_amt)
```

There is very little difference in median contribution amount between Clinton's
and Sanders' campaigns. The median contribution to Sanders' campaign is 27
dollars versus the median of 25 dollars for Clinton. However, the mean
contribution amount for Clinton is notably higher than that of Sanders: 103
dollars versus 41 dollars. The variance is much higher for Clinton's
contributions as well.

Next I'll write a function for making a histogram of a candidate's contributions
and use that for making histograms for individual candidates. The first two
candidates I'll use this function (`make_contb_hist`) for Clinton and Sanders,
the two leading Democratic candidates in the election.

```{r echo=FALSE, message=FALSE, warning=FALSE, Democratic_Cont_Histograms}
# Create function for plotting histograms with election data given x limits
# and binsize
make_contb_hist <- function(dataset, binsize, lower_limit, upper_limit) {
  hist <- qplot(x = contb_receipt_amt, data = dataset, binwidth = binsize) +
    scale_x_continuous(limits = c(lower_limit, upper_limit))
  return(hist)
}

# Use new function and grid.arrange (from gridExtra) to create histograms
clinton_hist <- make_contb_hist(clinton, 25, 0, 1000) +
  ggtitle("Clinton")

sanders_hist <- make_contb_hist(sanders, 25, 0, 1000) +
  ggtitle("Sanders")

grid.arrange(clinton_hist, sanders_hist, ncol = 2)
```

Both of these histograms are positively skewed and reflect what we're seeing
in the summaries. Considering that most contributions are 250 dollars or less
each, I'd like to zoom in a bit to get a better sense of the shape of these
distributions.

```{r echo=FALSE, message=FALSE, warning=FALSE, Dem_Cont_Hist_Zoom}
# Create two new zoomed-in histograms and display with grid.arrange
clinton_hist_zoom <- make_contb_hist(clinton, 5, 0, 250) +
  ggtitle("Clinton")

sanders_hist_zoom <- make_contb_hist(clinton, 5, 0, 250) +
  ggtitle("Sanders")

grid.arrange(clinton_hist_zoom, sanders_hist_zoom, ncol = 2)
```

As is the case for the overall data for receipt amounts, most of the
contributions to both Democratic candidates were 50 dollars or less. This seems
to make sense given that most contributors probably don't put aside very much
of their disposable income for presidential campaign contributions.

It's notable that the maximum refund and maximum contribution amounts to
Clinton's campaign are -2,700 and 2,700 dollars respectively. The maximum
contribution to Sanders' campaign is also 2,700 dollars, with the maximum refund
at the -2,500 mark. There aren't any outliers in this dataset that immediately
suggest contributions above the legal amount.

I'd like to see what the distribution of contributions for both Democratic
candidates looks like with a log10 transformation in light of this
transformation having been helpful with transforming the overall contribution
amount data.

```{r echo=FALSE, message=FALSE, warning=FALSE, Clinton_Sanders_Log_Transform}
# Create two histograms of Clinton & Sanders contributions with x-axis
# transformation using scale_x_log10. (Not using make_contb_hist because
# the function doesn't include an x-scale transformation variable.)
qplot(x = contb_receipt_amt, 
      data = subset(ga, cand_nm == 'Clinton, Hillary Rodham' 
                    | cand_nm == 'Sanders, Bernard')) +
  scale_x_log10() +
  facet_wrap(~cand_nm, ncol = 2)
```

The log10 transformation of these distributions appear to be approximately
normal, with a slight positive skew to the contributions received by Clinton
and a slight negative skew to contributions received by Sanders.

I'd like to see how the top Republican candidates compare in this regard.

Donald Trump:

```{r echo=FALSE, message=FALSE, warning=FALSE, Trump_Summary}
# Create subset of dataset containing contributions to Donald Trump
trump <- subset(ga, cand_nm == "Trump, Donald J.")
# Display summary of contribution amounts to Trump.
summary(trump$contb_receipt_amt)
```

Ted Cruz:

```{r echo=FALSE, message=FALSE, warning=FALSE, Cruz_Summary}
# Create subset of dataset containing contributions to Ted Cruz
cruz <- subset(ga, cand_nm == "Cruz, Rafael Edward 'Ted'")
# Display summary of contribution amounts to Cruz.
summary(cruz$contb_receipt_amt)
```

Ben Carson:

```{r echo=FALSE, message=FALSE, warning=FALSE, Carson_Summary}
# Create subset of dataset containing contributions to Ben Carson
carson <- subset(ga, cand_nm == "Carson, Benjamin S.")
# Display summary of contribution amounts to Carson.
summary(carson$contb_receipt_amt)
```

Jeb Bush:

```{r echo=FALSE, message=FALSE, warning=FALSE, Bush_Summary}
# Create subset of dataset containing contributions to Jeb Bush
bush <- subset(ga, cand_nm == "Bush, Jeb")
# Display summary of contribution amounts to Bush.
summary(bush$contb_receipt_amt)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Republican_Cont_Histograms}
# Create histograms and display them in 2x2 grid using grid.arrange
trump_hist <- make_contb_hist(trump, 25, 0, 1000) +
  ggtitle("Trump")

cruz_hist <- make_contb_hist(cruz, 25, 0, 1000) +
  ggtitle("Cruz")

carson_hist <- make_contb_hist(carson, 25, 0, 1000) +
  ggtitle("Carson")

bush_hist <- make_contb_hist(bush, 25, 0, 1000) +
  ggtitle("Bush")

grid.arrange(trump_hist, cruz_hist, carson_hist, bush_hist, ncol = 2)
```

As with the Democratic candidates, most contributions are less than 250 each.
I'd like to zoom in a bit and get a better sense of what the distribution of
Republican candidate contributions are in amounts of 250 dollars or less.

```{r echo=FALSE, message=FALSE, warning=FALSE, Republican_Cont_250}
# Create zoomed-in histograms of contributions < 250 for Republicans
trump_hist_zoom <- make_contb_hist(trump, 5, 0, 250) +
  ggtitle("Trump")
cruz_hist_zoom <- make_contb_hist(cruz, 5, 0, 250) +
  ggtitle("Cruz")
carson_hist_zoom <- make_contb_hist(carson, 5, 0, 250) +
  ggtitle("Carson")
bush_hist_zoom <- make_contb_hist(bush, 5, 0, 250) +
  ggtitle("Bush")

# Display new zoomed-in histograms in 2x2 grid using grid.arrange
grid.arrange(trump_hist_zoom, cruz_hist_zoom, carson_hist_zoom, 
             bush_hist_zoom, ncol = 2)
```

It's now clear that most of the contributions to the leading Republican
campaigns were less than 100 dollars. Only Trump's campaign has a significant
spike at the 200-dollar level.

I'd like to see what the Republican contribution data looks like after a log10
transformation on the x-axis.

```{r echo=FALSE, message=FALSE, warning=FALSE, Repub_Cont_Log_Transform}
# Create histograms of contributions to Republicans faceted by candidate name.
# Apply a log10 transformation to the x-axis on each histogram.
# (Here, and in certain later blocks of code, I forgo using make_contb_hist due
# to the axes transformations I want to try out.)
qplot(x = contb_receipt_amt, 
      data = subset(ga, cand_nm == 'Trump, Donald J.' 
                    | cand_nm == "Cruz, Rafael Edward 'Ted'" 
                    | cand_nm == 'Carson, Benjamin S.' 
                    | cand_nm == 'Bush, Jeb')) +
  scale_x_log10() +
  facet_wrap(~cand_nm, ncol = 2)
```

The log10-transformed distribution of contributions received by Bush is
relatively uniform. Carson, Cruz, and Trump received significantly more
contributions than Bush, and the log-transformed distribution of
contributions for these three are normal-ish, particularly those for Cruz and
Trump.

For the sake of completeness and comparison I'd like to look at the distribution
of contributions to third-party candidates.

```{r echo=FALSE, message=FALSE, warning=FALSE, Third_Party_Hist}
# Create histograms of contributions to third-party candidates faceted by name.
qplot(x = contb_receipt_amt, 
      data = subset(ga, cand_party != 'Democrat' 
                    & cand_party != 'Republican')) +
  facet_wrap(~cand_nm, ncol = 3)
```

There are magnitudes fewer contributions to these third-party candidates but the
data appears nonetheless long-tailed for Gary Johnson, the Libertarian nominee,
and Jill Stein, the Green nominee. Based on what I looked at with the previous
distributions, I'd like to see what a log10 transformation of the x-axis shows.

```{r echo=FALSE, message=FALSE, warning=FALSE, Third_Party_Hist_Log_Transform}
# Create histograms of third-party candidate contributions faceted by name.
# Apply a log10 scale transformation to the x-axis.
qplot(x = contb_receipt_amt, 
      data = subset(ga, cand_party != 'Democrat' 
                    & cand_party != 'Republican')) +
  scale_x_log10() +
  facet_wrap(~cand_nm, ncol = 3)
```

The contributions to Johnson and Stein have normal-ish distributions, while
McMullin's contribution distribution is comparatively uniform. McMullin only
received 41 contributions in Georgia so the uniformity of the transformed
distribution does not seem to convey any special meaning.

Taking a step back, it'd be handy to look at the two major-party nominees,
Trump and Clinton, side by side in both summaries and histogram form.

Trump:

```{r echo=FALSE, message=FALSE, warning=FALSE, Trump_Dist_Summary}
# Get a summary of contribution receipt amounts for Trump
summary(trump$contb_receipt_amt)
```

Clinton:

```{r echo=FALSE, message=FALSE, warning=FALSE, Clinton_Dist_Summary}
# Get a summary of contribution receipt amounts for Clinton
summary(clinton$contb_receipt_amt)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Nominee_Histograms}
# Create subset of data containing two major-party nominees
major_party_nominees <- subset(ga, cand_nm == "Trump, Donald J."
                               | cand_nm == "Clinton, Hillary Rodham")

# Plot contribution receipt amounts histogram on continuous scale
qplot(x = contb_receipt_amt,
      data = major_party_nominees,
      xlim = c(0, 1000)) +
  facet_wrap(~cand_nm, ncol = 2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Nominees_Log_Transform}
# Plot contribution receipt amounts histogram on log10 scale
qplot(x = contb_receipt_amt, 
      data = major_party_nominees) +
  scale_x_log10() +
  facet_wrap(~cand_nm, ncol = 2)
```

It's the Trump contribution distribution that contains the extreme outliers of
-5,900 dollars and 12,500 dollars. Presumably these amounts reflect
contributions that were divided up between the two elections - primary
and general - or contributions that were refunded at least in part in order to
remain under the 2,700-dollar individual threshold. 

Trump had higher mean and median contribution amounts, but Clinton had a higher
number of contributions overall, as is evident in the histograms. Both
distributions appear approximately normal after a log10 transformation of the
x-axis.

## Contribution Counts by Candidate and Date
I'd like to look at the dates that Clinton and Trump received contributions.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contrib_Dates_Clinton_Trump}
# Create histograms of contribution counts by date for major party candidates
# faceted by candidate name.
qplot(x = contb_receipt_dt, 
      data = major_party_nominees) +
  scale_x_date() +
  facet_wrap(~cand_nm)
```

The shapes of these graphs are really interesting. Clinton's contributions
steadily climb for the most part, with a dip occuring after the month of August
2016. Trump's contributions go from miniscule to quite a lot as July arrives,
after which there's a dip in October (not unlike Clinton's dip in August) which
precedes upticks of contributions right before and after the election.

There's an uptick in July for both candidates, which may have something to do
with the Democratic and Republican National Conventions taking place in the
month of July. These conventions mark when these two candidates officially
became the major party nominees.

I'd like to get a closer look at the last three months of 2016. The election
occurred on November 8, 2016, but there are contributions marked as late as
December 31, 2016. I'm going to zoom in on these last three months and adjust
the binwidth to display individual days.

```{r echo=FALSE, message=FALSE, warning=FALSE, Oct_Dec_Clinton_Trump}
# Create a histogram of contribution counts between 10/1/2016 and 12/31/2016 by
# date to the major-party nominees, faceted by candidate name.
qplot(x = contb_receipt_dt, 
      data = subset(ga, cand_nm == 'Clinton, Hillary Rodham' | 
                      cand_nm == 'Trump, Donald J.'), binwidth = 1) +
  scale_x_date(limits = c(as.Date('2016-10-01'), as.Date('2016-12-31'))) +
  facet_wrap(~cand_nm)
```

Contributions to both campaigns climbed and spiked in mid-October before dipping
slightly. From there the contributions to Clinton's campaign ramped up again,
peaking in the days right before the election.

Trump continued to receive contributions after Election Day. These contributions
steadily trickled in through early December.

Looking at a frequency polygon of this data might be useful.

```{r echo=FALSE, message=FALSE, warning=FALSE, Clinton_Trump_FreqPoly}
# Create frequency polygons for contribution counts by date to major party
# nominees, colored by candidate name.
qplot(x = contb_receipt_dt, 
      data = major_party_nominees,
      binwidth = 1, geom = 'freqpoly', colour = cand_nm) +
  scale_x_date(limits = c(as.Date('2016-10-01'), as.Date('2016-12-31')))
```

These frequency polygons provide a nice chart of the last couple of months of
the general election and how contributions to the two major party nominees
unfolded during that time. Clinton's fundraising in Georgia was considerably
higher than that of Trump up until the moment of the election, after which Trump
continued to receive contributions through the end of the calendar year.

Here's a version of the figure above that encompasses the entire general
election period after Clinton and Trump were formally nominated as their
party's nominees.

```{r echo=FALSE, message=FALSE, warning=FALSE, General_Election_FreqPoly}
# Create frequency polygons for contribution counts by date to major party
# nominees, colored by candidate name, for the entire general election period.
qplot(x = contb_receipt_dt, 
      data = major_party_nominees,
      binwidth = 1, geom = 'freqpoly', colour = cand_nm) +
  scale_x_date(limits = c(as.Date('2016-07-01'), as.Date('2016-12-31')),
               date_breaks = "1 month")
```

In the multivariate plots section, I'd like to look at contribution totals by
date. This seems like it'd be helpful to get a more full picture of how campaign
contributions changed over the course of the election.

## Contribution Amounts vs. Date
I'd like to look at a scatterplot showing individual contribution amounts over
the course of time. I'll start by taking a look at a scatterplot of every single
transaction, positive and negative, over the range of dates represented in the
dataset.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_All}
# Create scatterplot of contribution amounts by date
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt), data = ga) +
  geom_point()
```

There is some serious overplotting going on here, and outliers in multiple
directions. I'd like to zoom in on the area between 0 and 2700 dollars (the
upper limit on individual contributions per election), as well as the time
period from early 2015 through the end of 2016, where the vast majority of
contributions appear to have taken place. I'll also use the `alpha` parameter
for `geom_point` to add some transparency and hopefully reduce the effect of the
overplotting.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_All_2700}
# Create scatterplot of contribution amounts by date with 1/5 transparency,
# from 3/1/2015-12/31/2016, under the amount of 2,700 dollars.
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt), data = ga) +
  geom_point(alpha = 0.2) +
  scale_x_date(limits = c(as.Date('2015-03-01'), as.Date('2016-12-31'))) +
  coord_cartesian(ylim = c(0, 2700))
```

Given my earlier explorations, it's no surprise that most contributions are
under 500 dollars apiece. I'd like to zoom in even more and increase the
transparency to see if that tells us anything new.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_All_500}
# Create scatterplot of contribution amounts by date with 1/100 transparency,
# from 3/1/2015-12/31/2016, under the amount of 500 dollars.
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt), data = ga) +
  geom_point(alpha = 0.01) +
  scale_x_date(limits = c(as.Date('2015-03-01'), as.Date('2016-12-31'))) +
  coord_cartesian(ylim = c(0, 500))
```

Even with transparency reduced to 1/100, we're still seeing massive overplotting
between the amounts of 0 and 50 dollars. I'll zoom in once more and introduce
some noise to the plot using `geom_jitter` to see what that shows.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_All_50}
# Create scatterplot of contribution amounts by date with 1/100 transparency,
# from 3/1/2015-12/31/2016, under the amount of 50 dollars.
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt), data = ga) +
  geom_jitter(alpha = 0.01) +
  scale_x_date(limits = c(as.Date('2015-03-01'), as.Date('2016-12-31'))) +
  coord_cartesian(ylim = c(0, 50))
```

This gives a better picture of the frequency of particular contribution levels
over time. Starting in late 2015, contributions at the 25- and 50-dollar
levels ramp up at what appears to be an almost identical rate and do not let
up until early November 2016. Most contribution amounts are multiples of 5. The
most solid-looking (but not lengthiest) lines observable in the graph are
at the 5- and 10-dollar levels. The lengthiest lines are at the 25- and
50-dollar levels.

## Individual Contributor Totals
If possible I'd like to take a stab at looking at total contributions from
individual contributors in Georgia. In order to do this I'm going to use the
`group_by`, `select`, and `summarise` functions from the `dplyr` package to try
and boil down the data.frame to individual contributor names and their total
contributions to different candidates (allowing for the possibility that
they may have donated to more than one candidate).

```{r echo=FALSE, message=FALSE, warning=FALSE, Indiv_Contrib_Reshape}
indiv_columns <- select(ga, contbr_nm, contbr_city, contbr_zip, contbr_employer,
                       contbr_occupation, cand_nm, contb_receipt_amt)
indiv_columns$contbr_zip <- as.numeric(substr(indiv_columns$contbr_zip, 0, 5))

indiv_totals <- indiv_columns %>% 
  group_by(contbr_nm, contbr_city, contbr_zip, contbr_employer,
                          contbr_occupation, cand_nm) %>% 
  summarise(sumcontb = sum(contb_receipt_amt))
```

Here are the first 20 rows of a data.frame in which I try to consolidate
individual contributor totals into a row for each contributor and the total
they gave to a particular candidate:

```{r echo=FALSE, message=FALSE, warning=FALSE, Head_Individual_Totals}
# Look at first 20 rows of data.frame containing attempt at individual total
head(indiv_totals, 20)
```

There are a few problems that make it challenging to do a proper analysis of
individual contribution totals. The primary issue is that it's difficult to
definitively consolidate individual contributors into rows that accurately
reflect the total they've contributed to particular candidates.

The reason for this problem is not that the information is inaccurate but that
contributors may have given different answers to the same demographic questions
with different contributions they gave. For example, the name David Aaron
appears in rows 8 and 9 of the data.frame excerpted above. Both rows have the
same name, city, zip code, profession, and candidate (Trump) but different
employers are specified. Is this one person who changed employers at some point
during the election? Or do these rows represent a father and son with the same
name and profession? It's hard to say for sure.

# Form Type vs. Candidate Party
I'm curious if there's any relationship between party and form type.

```{r echo=FALSE, message=FALSE, warning=FALSE, Form_Types_by_Party}
# Display summary of form types by candidates' political parties
by(ga$form_tp, ga$cand_party, summary)
```

There were almost twice as many contributions to Republican candidates from
"authorized committees" than there were to Democratic candidates. Meanwhile
nearly all of the contributions to third-party candidates were from individuals.

In the multivariate plots section I'll take a closer look at whether there are
any relationships between form type and contribution amount at different times
in the election.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
One relationship I observed is between date and the number of contributions to
major-party candidates. In particular Trump's contributions from Georgia
spiked considerably around July 2016, the month of both the Republican
and Democratic National Conventions.

Another relationship I observed is between party and contribution amount.
Republican candidates received considerably higher contribution dollar amounts 
on average versus Democratic candidates. I thought this was particularly
notable given the high number of Republican candidates who were in the running
early in the election.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
There were almost twice many "authorized committee" contributions to Republican
candidates than there were to Democratic candidates. 

Additionally, all but one contribution to third-party nominees were made by
individuals and not by committees.

I also realized that, due to the wide range of potentially correct answers (as
of the time of a contribution) an individual contributor could specify different
demographic answers each time and it'd still be good data, even if it's hard to
organize by individual donor.

### What was the strongest relationship you found?
The strongest relationship I came across is the relationship between candidate
party and contribution amount. For Trump and all the other Republican
candidates, the mean and median contribution levels were much higher than they
were for Democratic candidates, despite the glut of Republican candidates
early on in the elections.

# Multivariate Plots Section

## A Closer Look at Contributions Over Time
Expanding on my bivariate explorations, I'd like to look at how the individual
candidates compared to each other as far as the contributions they've received
over time.

I'll look at Trump's contribution amounts over time at first, starting with a
scatterplot that shows all the contributions and refunds, including all the
outliers, alongside the previous graph.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_All_Trump}
# Create scatterplot of contribution amounts to Trump by date
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt), data = trump) +
  geom_point() +
  ggtitle("Contributions to Trump")
```

This is kind of ugly and overplotted, but it at least initially looks quite
similar to the first graph I made of contributions over time. I'd like to
compare this distribution with the plot of all contributions. In light of the
earliest Trump data going back as far as late 2014, I'll put this new
plot and the original plot on the same scale to get a sense of the change over
time.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_Trump_V_All}
# Create plot for all contributions from 12/2014 through 12/2016
all_contbs_time <- ggplot(aes(x = contb_receipt_dt,
                              y = contb_receipt_amt), data = ga) +
  geom_point() +
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  ggtitle("All Contributions Over Time (Dec. '14 - Dec. '16)")

# Create plot for all contributions to Trump 12/2014 - 12/2016
trump_contbs_time <- ggplot(aes(x = contb_receipt_dt,
                                y = contb_receipt_amt), data = trump) +
  geom_point() +
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  ggtitle("Contributions to Donald Trump Over Time (Dec. '14 - Dec. '16)")

# Display both plots create above in one column
grid.arrange(all_contbs_time, trump_contbs_time, ncol = 1)
```

The set of contributions to all candidates goes back as far as the very tail end
of 2014 and the first contributions to Trump's primary campaign start showing up
in June 2015, and take a massive uptick shortly before July 2016. As mentioned
earlier, this is when the Republican and Democratic Conventions were held. 

Having these two graphs on the same y-scale and the same x-axis length reaffirms
that a lot of the highest and lowest contribution amounts in the dataset are
accounted for in the contributions to Trump.

Let's look at the contributions over time to Clinton's campaign and then compare
them with the two scatterplots we've been looking at.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_Clinton}
# Plot all contributions to Clinton over time on Date scale
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt), data = clinton) +
  geom_point() +
  ggtitle("Contributions to Clinton")
```

Once again there's significant overplotting here, and, due to its variance, this
data also shows up on a different y-scale than the one we saw looking at the
state's contributions to Trump's campaign. The earlier summaries showed that all
of the reported individual contributions to Clinton's campaign fall between
-2,700 and 2,700 dollars. The vast majority of contributions appear to be in the
1- to 500-dollar range.

Before zooming in to take a closer look, I'll take a quick glance at how the
previous two scatterplots compare with that of Clinton when all three are
plotted on the same y-scale, positive and negative outliers included.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_T_vs_C_vs_All}
# Create scatterplot for all contributions to Clinton from 12/2014 to 12/2016
clinton_contbs_time <- ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt),
                              data = clinton) +
  geom_point() +
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  ylim(-5900, 12500) +
  ggtitle("Contributions to Hillary Clinton Over Time (Dec. '14 - Dec. '16)")

# Display Clinton scatterplot in one column below all contribs / Trump contribs
grid.arrange(all_contbs_time, trump_contbs_time, clinton_contbs_time, ncol = 1)
```

The contributions to Clinton are more overplotted than those received by Trump
and are visibly higher in sheer quantity and in consistency of occurrence over
time, even if the variance is not as high as that of Trump. This graph certainly
makes it look like Clinton's contributions comprise a big percentage of the
overall contributions, but the overplotting makes it hard to tell what's really
going on.

I'm going to put all three of these plots on a y-scale that covers the range of
contribution amounts and refunds that's visible in Clinton's scatterplot. I'll
also add some transparency and noise in order to be able to look at them in a
way that's useful. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_T_vs_C_vs_All_Adj}
# Save previous plots to variables and plot on same y-scale (0 to 500)
all_contbs_time_adj <- ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt),
                              data = ga) +
  geom_jitter(alpha = 0.01) +
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  ylim(0, 500) +
  ggtitle("All Contributions Over Time (Dec. '14 - Dec. '16)")

trump_contbs_time_adj <- ggplot(aes(x = contb_receipt_dt, 
                                    y = contb_receipt_amt), data = trump) +
  geom_jitter(alpha = 0.01) +
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  ylim(0, 500) +
  ggtitle("Contributions to Donald Trump Over Time (Dec. '14 - Dec. '16)")

clinton_contbs_time_adj <- ggplot(aes(x = contb_receipt_dt, 
                                      y = contb_receipt_amt), data = clinton) +
  geom_jitter(alpha = 0.01) +
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  ylim(0, 500) +
  ggtitle("Contributions to Hillary Clinton Over Time (Dec. '14 - Dec. '16)")

grid.arrange(all_contbs_time_adj, trump_contbs_time_adj,
             clinton_contbs_time_adj, ncol = 1)
```

It looks like both Clinton and Trump were drawing in a significant amount of
contributions in the range of 100 to 300 dollars during the months between the
party conventions and the election, but Clinton had a more consistent amount of
contributions in the 1- to 100-dollar range up until the election.

After Trump won, a small amount of contributions came in before 2016's end, many
of which were around the 100-dollar amount.

In light of my discovery that the vast majority of contributions were 500
dollars or less, I'd like to look at a scatterplot of contributions in the 1- to
500-dollar range, and impose a `geom_line` over the scatterplot showing median
contribution levels, using a different color for each group.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_Adj_500}
# Create scatterplot showing contributions over time between 0 and 500 dollars
# Save to variable
all_contbs_time_adj_500 <- ggplot(aes(x = contb_receipt_dt,
                                      y = contb_receipt_amt), data = ga) +
  # Set transparency to 1/100 and jitter points on scatterplot
  geom_jitter(alpha = 0.01) +
  # Add orange line representing median contribution amount
  geom_line(stat = 'summary', fun.y = median, colour = 'orange') +
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  coord_cartesian(ylim = c(0, 500)) +
  ggtitle("All Contributions (< $500) (Dec. '14 - Dec. '16)")

# Create plot showing contributions to Trump over time between 0 and 500 dollars
# Save to variable
trump_contbs_time_adj_500 <- ggplot(aes(x = contb_receipt_dt,
                                        y = contb_receipt_amt), data = trump) +
  # Set transparency to 1/100 and jitter points on scatterplot
  geom_jitter(alpha = 0.01) +
  # Add red line representing median contribution amount
  geom_line(stat = 'summary', fun.y = median, colour = 'red') +
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  coord_cartesian(ylim = c(0, 500)) +
  ggtitle("Contributions (<$500) to Donald Trump (Dec. '14 - Dec. '16)")

# Create plot showing contributions to Clinton over time between 0 and 500 
# dollars. Save to variable.
clinton_contbs_time_adj_500 <- ggplot(aes(x = contb_receipt_dt,
                                          y = contb_receipt_amt),
                                      data = clinton) +
  # Set transparency to 1/100 and jitter points on scatterplot
  geom_jitter(alpha = 0.01) +
  # Add blue line representing median contribution amount
  geom_line(stat = 'summary', fun.y = median, colour = 'blue') +
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  coord_cartesian(ylim = c(0, 500)) +
  ggtitle("Contributions (<$500) to Hillary Clinton (Dec. '14 - Dec. '16)")

# Plot all three scatterplots together in one column
grid.arrange(all_contbs_time_adj_500, trump_contbs_time_adj_500,
             clinton_contbs_time_adj_500, ncol = 1)
```

These median lines are really quite noisy so I'll also try looking at this data
using `geom_smooth` to plot conditional means.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_Smooth_500}
# Create scatterplot showing contributions over time between 0 and 500 dollars
# Save to variable
all_contbs_time_smooth_500 <- ggplot(aes(x = contb_receipt_dt,
                                      y = contb_receipt_amt), data = ga) +
  # Set transparency to 1/100 and jitter points on scatterplot
  geom_jitter(alpha = 0.01) +
  # Add orange smoother line
  geom_smooth(colour = 'orange') +
  # 
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  coord_cartesian(ylim = c(0, 500)) +
  ggtitle("All Contributions (< $500) (Dec. '14 - Dec. '16)")

# Create plot showing contributions to Trump over time between 0 and 500 dollars
# Save to variable
trump_contbs_time_smooth_500 <- ggplot(aes(x = contb_receipt_dt,
                                        y = contb_receipt_amt), data = trump) +
  # Set transparency to 1/100 and jitter points on scatterplot
  geom_jitter(alpha = 0.01) +
  # Add red smoother line
  geom_smooth(colour = 'red') +
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  coord_cartesian(ylim = c(0, 500)) +
  ggtitle("Contributions (<$500) to Donald Trump (Dec. '14 - Dec. '16)")

# Create plot showing contributions to Clinton over time between 0 and 500 
# dollars. Save to variable.
clinton_contbs_time_smooth_500 <- ggplot(aes(x = contb_receipt_dt,
                                          y = contb_receipt_amt),
                                      data = clinton) +
  # Set transparency to 1/100 and jitter points on scatterplot
  geom_jitter(alpha = 0.01) +
  # Add blue smoother line
  geom_smooth(colour = 'blue') +
  scale_x_date(limits = c(as.Date('2014-12-01'), as.Date('2016-12-31'))) +
  coord_cartesian(ylim = c(0, 500)) +
  ggtitle("Contributions (<$500) to Hillary Clinton (Dec. '14 - Dec. '16)")

# Plot all three scatterplots together in one column
grid.arrange(all_contbs_time_smooth_500, trump_contbs_time_smooth_500,
             clinton_contbs_time_smooth_500, ncol = 1)
```

It'd take some careful parsing to attribute particular events to the various
peaks and dips in these graphs, but there are some noticeable trends going on
here. It looks like although Trump got a significant amount of contributions
from Georgia around the Republican National Convention in July, the number of
contributions he received later on was for the most part a trickle compared to
the amount of contributions coming in for Clinton. 

However, the median and median contribution amounts are, for much of election
season, much higher for Trump than they were for Clinton. He may not have had
the same number of supporters in Georgia as Clinton did, but the ones who
pledged money to his campaign gave bigger sums of money on average. There is a
significant jump up in mean contribution amount for Trump in the couple of
months after his nomination.

It's notable that the time periods in which there are a high number of
contributions coincide with a lower median contribution amount. This seems to
make sense in light of the fact that most people are unlikely to be able to give
very high contributions to presidential candidates, even if they want to.

I'm going to zero in on the period of the general election - July through
November of 2016. It'd be interesting to see what donations under $500 looked
like over time during this period. In addition to a scatterplot showing all the
contributions, as well as scatterplots of the contributions received by Clinton
and Trump, I'll make a fourth scatterplot showing contributions received by
third-party candidates during the general election.

```{r echo=FALSE, message=FALSE, warning=FALSE, Rcpt_Dt_Amt_T_vs_C_vs_All_Gen}
# Create scatterplot showing all contributions during general election period
# Save to variable
all_contbs_time_general <- ggplot(aes(x = contb_receipt_dt,
                                      y = contb_receipt_amt), data = ga) +
  geom_jitter(alpha = 0.01) +
  scale_x_date(limits = c(as.Date('2016-07-01'), as.Date('2016-12-31')),
               breaks = seq(as.Date('2016-07-01'), as.Date('2016-12-31'),
                            'month')) +
  ylim(0, 500) +
  ggtitle("All Contributions (< $500) Over Time (Jul. '16 - Dec. '16)")

# Create plot showing all contributions to Trump during general election
# Save to variable
trump_contbs_time_general <- ggplot(aes(x = contb_receipt_dt,
                                        y = contb_receipt_amt),
                                    data = trump) +
  geom_jitter(alpha = 0.01) +
  scale_x_date(limits = c(as.Date('2016-07-01'), as.Date('2016-12-31')),
               breaks = seq(as.Date('2016-07-01'), as.Date('2016-12-31'),
                            'month')) +
  ylim(0, 500) +
  ggtitle("Contributions (<$500) to Donald Trump (Jul. '16 - Dec. '16)")

# Create plot showing all contributions to Clinton during general election
# Save to variable
clinton_contbs_time_general <- ggplot(aes(x = contb_receipt_dt,
                                          y = contb_receipt_amt),
                                  data = clinton) +
  geom_jitter(alpha = 0.01) +
  scale_x_date(limits = c(as.Date('2016-07-01'), as.Date('2016-12-31')),
               breaks = seq(as.Date('2016-07-01'), as.Date('2016-12-31'),
                            'month')) +
  ylim(0, 500) +
  ggtitle("Contributions (<$500) to Hillary Clinton (Jul. '16 - Dec. '16)")

# Create plot showing all contributions to third-party candidates during general
# Save to variable
third_party <- subset(ga, cand_party != 'Democrat'
                      & cand_party != 'Republican')

third_party_contbs_time_gen <- ggplot(aes(x = contb_receipt_dt,
                                          y = contb_receipt_amt),
                                      data = third_party) +
  geom_jitter(alpha = 0.01) +
  scale_x_date(limits = c(as.Date('2016-07-01'), as.Date('2016-12-31')),
               breaks = seq(as.Date('2016-07-01'), as.Date('2016-12-31'),
                            'month')) +
  ylim(0, 500) +
  ggtitle("Contributions (<$500) to 3rd Party Candidates (Jul. '16 - Dec. '16)")

# Plot all four scatterplots together in one column
grid.arrange(all_contbs_time_general, trump_contbs_time_general,
             clinton_contbs_time_general, third_party_contbs_time_gen, ncol = 1)
```

The bulk of the contributions received during the general election clearly went
to Clinton, whose solid black streak under the $100 level through election day
represents the most consistent intake over time. With the level of transparency
here, the contributions to third-party candidates is negligible, with the
darkest area showing during the month of July (and not a very dark area at
that). 

## Debates and Campaign Turning Points

I do wonder about the ebb and flow of contributions to Trump in Georgia over the
course of this timespan, particularly because he won the state despite being
out-fundraised. Perhaps the presidential debates had some kind of effect on the
number of contributions that came in. I'll insert some vertical dotted lines
representing the presidential debate dates - 9/26 (blue), 10/9 (purple), and
10/19 (red).

```{r echo=FALSE, message=FALSE, warning=FALSE, General_Election_and_Debates}
# Add vertical lines to Trump and Clinton scatterplots representing debates

# Create new scatterplot containing general election contributions to Trump
trump_contbs_time_gen_debates <- trump_contbs_time_general +
  # Add blue dashed line representing September 26 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-09-26'))), linetype = 4,
             colour = 'blue') +
  # Add purple dashed line representing October 9 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-09'))), linetype = 4,
             colour = 'purple') +
  # Add red dashed line representing October 19 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-19'))), linetype = 4,
             colour = 'red')

# Create new scatterplot containing general election contributions to Clinton
clinton_contbs_time_gen_debates <- clinton_contbs_time_general +
  # Add blue dashed line representing September 26 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-09-26'))), linetype = 4,
             colour = 'blue') +
  # Add purple dashed line representing October 9 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-09'))), linetype = 4,
             colour = 'purple') +
  # Add red dashed line representing October 19 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-19'))), linetype = 4,
             colour = 'red')

# Display these new scatterplots together in a single column
grid.arrange(trump_contbs_time_gen_debates, clinton_contbs_time_gen_debates,
             ncol = 1)
```

There appear to be very slight bumps up in contributions for both Clinton and
Trump after the first debate on September 26th. This is noticeable in the
streaks that show up at the 100-, 200-, 250-, and 400-dollar levels for Trump,
and at the 100- and 250-dollar levels for Clinton in addition to her already
sustained levels of contributions of 50 dollars or less.

It's more difficult to notice any change after the second debate on October 9th.
There may be a very slight bump to contributions to Trump at the 100-, 200-,
and 250-dollar levels. The third debate seems to coincide with darkened sections
of the scatterplot showing up at the 200-, 250-, and 400-dollar levels for
Trump.

There were several dramatic news stories during the election that seemed to be
turning points. I'm curious whether the tape that surfaced of Trump's lewd
conversation with Billy Bush (released 10/7), as well as FBI director James
Comey's letter to the U.S. Congress (released 10/28) had any noticeable effect
on the number and amount of contributions coming in to the major candidates.

```{r echo=FALSE, message=FALSE, warning=FALSE, Tru_Clint_General_Contbs_Events}
# Add two election turning points to scatterplots of general election
# contributions to Trump and Clinton

# Create scatterplot showing contributions to Trump during general election
trump_contbs_time_gen_events <- trump_contbs_time_general +
  # Add vertical green dashed line representing date of release of Bush tape
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-07'))), linetype = 4,
             colour = 'green') +
  # Add vertical orange dashed line representing date of Comey letter
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-28'))), linetype = 4,
             colour = 'orange')

# Create scatterplot showing contributions to Clinton during general election
clinton_contbs_time_gen_events <- clinton_contbs_time_general +
  # Add vertical green dashed line representing date of release of Bush tape
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-07'))), linetype = 4,
             colour = 'green') +
  # Add vertical orange dashed line representing date of Comey letter
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-28'))), linetype = 4,
             colour = 'orange')

# Display these new scatterplots together in one column
grid.arrange(trump_contbs_time_gen_events, clinton_contbs_time_gen_events,
             ncol = 1)
```

I'm having a harder time noticing any effects here - neither of these vertical
lines representing the dates in question appear to correspond with any major
changes in the scatterplot. There may be a very slight increase in contributions
to Clinton after the Trump-Billy Bush tape, but it's hard to tell. That's not to
say that these events didn't have an effect on the election - the Comey letter
in particular is [widely argued](https://fivethirtyeight.com/features/the-comey-letter-probably-cost-clinton-the-election/) to have cost Clinton the election. Effects on
the number of contributions and the amounts of these contributions are more
difficult to pick out in these scatterplots.

## Contribution Totals by Day
The next set of multivariate plots I'd be interested in looking at are total
contributions in dollars received each day by each candidate over the course of
the primary and general elections in 2016.

What I'll do first is use `dplyr` to select and reshape the data in order to
get daily total contributions for every candidate in a R data.frame. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Contb_Totals_by_Day_Dplyr}
# Select particular columns from dataset: candidate name, contribution receipt
# date, contribution receipt amount, and transaction ID.
daily_totals_columns <- select(ga, cand_nm, contb_receipt_dt, contb_receipt_amt,
                               tran_id)
# Group the new data.frame by candidate name and receipt date, and calculate
# daily totals to add as a column to these new in daily_totals data.frame
daily_totals <- daily_totals_columns %>% 
  group_by(cand_nm, contb_receipt_dt) %>% 
  summarise(daily_total = sum(contb_receipt_amt))
```

I'll do a set of line graphs on one set of axes using `ggplot2`. The color of
each line in the plot represents an individual candidate.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contb_Totals_by_Day_LinePlot}
# Create line graphs of daily contribution totals received, colored by candidate
ggplot(aes(x = contb_receipt_dt, y = daily_total, colour = cand_nm),
       data = daily_totals) +
  geom_line()
```

This map is colorful but also very noisy. It's easy to see the pink line on the
right end of the graph representing Trump, which has a huge spike in the middle
and has a lot of noisy variance approaching the election. More colors are
perceptible on the left side of the graph when there were still a lot of
Republican candidates contending for the nomination. 

It may be helpful to zoom in on this a bit. I'll look at 2016 only and see if
this imparts anything useful.

```{r echo=FALSE, message=FALSE, warning=FALSE, Daily_Totals_Zoom}
# Create line graphs of daily contribution totals received, colored by candidate
ggplot(aes(x = contb_receipt_dt, y = daily_total, colour = cand_nm),
       data = daily_totals) +
  geom_line() +
  # Zoom in to 2016 and contributions between 0 and 5,400 dollars
  coord_cartesian(xlim = c(as.Date('2016-01-01'), as.Date('2016-12-31')),
                  ylim = c(0, 5400))
```

As colorful as this is, it's an incredibly noisy graph and is kind of hard to
glean information from. 

Perhaps looking at daily contribution totals by party would be useful. First I'll
use `dplyr` to create a new data.frame containing daily contribution totals by
party.

```{r echo=FALSE, message=FALSE, warning=FALSE, Daily_Party_Contb_Totals}
# Select columns from dataset: party, date, amount, transaction ID
daily_party_totals_cols <- select(ga, cand_party, contb_receipt_dt, 
                                  contb_receipt_amt, tran_id)
# Group columns in new data.frame by party and receipt date
# Add new column for sum daily total contributions received by party
daily_party_totals <- daily_party_totals_cols %>% 
  group_by(cand_party, contb_receipt_dt) %>% 
  summarise(daily_party_total = sum(contb_receipt_amt))
```

Next I'll plot contribution totals by day in a line graph, with the line colors
reflecting political party.

```{r echo=FALSE, message=FALSE, warning=FALSE, Daily_Party_Contb_Totals_Graph}
# Plot line graphs for 2016 showing contributions received by candidate party
ggplot(aes(x = contb_receipt_dt, y = daily_party_total, colour = cand_party),
       data = daily_party_totals) +
  geom_line() +
  scale_x_date(lim = c(as.Date('2016-01-01'), as.Date('2016-12-31')))
```

This is much easier to read than the graph showing every one of the candidates.
It appears that the most successful fundraising parties day by day were the
Democrats and the Republicans, with both of those parties intermittently
passing up each other as far as daily total contributions in dollars. The
variance in Trump's daily totals is much higher than that of Clinton, but he
still appeared to out-fundraise her on at least a few days during the election.

I'd also like to take a look at daily totals by candidate gender. I'll use
`dplyr` to make a data.frame that reflections daily totals by candidate gender.

```{r echo=FALSE, message=FALSE, warning=FALSE, Daily_Gender_Totals}
# Select columns from ga data.frame to create new data.frame: gender,
# contribution receipt amount, receipt date, and transaction ID.
daily_totals_gender_cols <- select(ga, cand_gender, contb_receipt_dt, 
                                   contb_receipt_amt, tran_id)
# Group new data.frame by gender and date, and add new column with daily total
daily_totals_gender <- daily_totals_gender_cols %>% 
  group_by(cand_gender, contb_receipt_dt) %>% 
  summarise(daily_total_gender = sum(contb_receipt_amt))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Daily_Gender_Totals_Graph}
# Plot new daily totals by gender data.frame for year 2016
# Set upper & lower limits to 0 and 125000 dollars
ggplot(aes(x = contb_receipt_dt, y = daily_total_gender, colour = cand_gender),
       data = daily_totals_gender) +
  geom_line() +
  coord_cartesian(xlim = c(as.Date('2016-01-01'), as.Date('2016-12-31')),
                  ylim = c(0, 125000))
```

It appears that for most of 2016 there were more high daily dollar-amounts going
to male candidates than female candidates. There are just a few days in the
first half of the year in which female candidates out-fundraise the male
candidates. The fact that female candidates managed to be ahead at this stage,
when there are so many male competitors, seems noteworthy.

### Daily Totals vs. Election Turning Points
I'd like to see if those turning points I looked at a bit earlier have any sort
of discernible pattern atop the graph of contributions by party.

First I'll add dashed vertical lines representing the debates to the graphs of
daily totals by party.

```{r echo=FALSE, message=FALSE, warning=FALSE, Daily_Totals_Party_Debates}
# Plot line graphs for 2016 showing contributions received by candidate party
ggplot(aes(x = contb_receipt_dt, y = daily_party_total, colour = cand_party),
       data = daily_party_totals) +
  geom_line() +
  # Add blue dashed line representing September 26 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-09-26'))), linetype = 4,
             colour = 'blue') +
  # Add purple dashed line representing October 9 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-09'))), linetype = 4,
             colour = 'purple') +
  # Add red dashed line representing October 19 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-19'))), linetype = 4,
             colour = 'red') +
  # Restrict x-axis to September 1 through November 15
  # Restrict y-axis to contributions between 0 to 100000 without removing values
  coord_cartesian(xlim = c(as.Date('2016-09-01'), as.Date('2016-11-15')),
                  ylim = c(0, 100000))
```

These line graphs give a much different look at the potential effects of the
presidential debates on campaign contributions from Georgia.

All three debates appear to mark points where Georgia's contributions to the
Democratic candidate (Clinton) dropped and contributions to the Republican
candidate (Trump) increased. There's all sorts of up-and-down activity before,
in between, and after the debates, but all of a sudden it looks like the debates
helped Trump and hurt Clinton when it came to Georgia contributions.

There doesn't appear to be a whole lot of connection between the debates and
the contributions hauled in by the third-party candidates. This would seem to
make sense in light of the third-party candidates not being invited to take part
in the televised debates. However, one of the main spikes in contributions to
the Libertarian candidate, Gary Johnson, shows up shortly after the first
debate. Perhaps right-leaning voters in the state took an increased interest in
Johnson after the first debate.

Next I'll turn to the aforementioned turning point events - the Billy Bush tape
and James Comey's letter to Congress. I'll try laying lines representing those
events over the graphs of daily totals by party.

```{r echo=FALSE, message=FALSE, warning=FALSE, Daily_Totals_Party_Events}
ggplot(aes(x = contb_receipt_dt, y = daily_party_total, colour = cand_party),
       data = daily_party_totals) +
  geom_line() +
  # Add vertical green dashed line representing Billy Bush tape release
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-07'))), linetype = 4,
             colour = 'green') +
  # Add vertical orange dashed line representing date of Comey letter
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-28'))), linetype = 4,
             colour = 'orange') +
  # Restrict x-axis to September 1 through November 15
  # Restrict y-axis to contributions between 0 to 100000 without removing values
  coord_cartesian(xlim = c(as.Date('2016-09-01'), as.Date('2016-11-15')),
                  ylim = c(0, 100000))
```

There was a drop in contributions to both Trump and Clinton right after the
release of the Billy Bush tape. Additionally, there appears to be a drop in
contributions for Trump right after the James Comey letter while there's a brief
increase in contribution totals for Clinton. These outcomes are rather different
than I expected - I thought that the Billy Bush tape would help Clinton and hurt
only Trump, and thought that the Comey letter would help Trump and hurt only
Clinton. It looks like, in this particular case, my expected outcomes were
incorrect.

## Election Type vs. Contribution Amount
I am curious whether contribution amounts are affected by the type of election
taking place, whether it's the primary or general election. First I'll use the
`by` function to take a look at summaries of contribution amounts in each
election.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contributions_by_Election_Type}
# Display summaries of contribution amounts by election type
by(ga$contb_receipt_amt, ga$election_tp, summary)
```

The mean and median contribution amounts for primary elections (labeled as
`P2016` in the `election_tp` column) are only slightly lower than the mean and
median for general elections (labeled as `G2016` in the `election_tp` column).
At first glance, it doesn't appear that there's a huge difference in overall
contribution amounts between the 2016 primary and general elections.

I'd like to look more specifically at summaries of contribution amounts to
the two major-party candidates in each election.

```{r echo=FALSE, message=FALSE, warning=FALSE, Trump_Contbs_by_Type_Summary}
by(trump$contb_receipt_amt, trump$election_tp, summary)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Clinton_Contbs_by_Type_Summary}
by(clinton$contb_receipt_amt, clinton$election_tp, summary)
```

Trump's median and mean contribution amounts nearly doubled once the general
election was underway. Clinton's median contribution amount, meanwhile, was
exactly the same for the primary and general - 25 dollars - and the mean
contribution amount went down during the general election from 128.2 dollars
to 88.59 dollars.

In light of my earlier finding that around 450 transactions have an `election_tp`
of "NA" I'll take a look at a summary of contribution amounts labeled this way.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contributions_Type_NA_2016}
summary(na_2016$contb_receipt_amt)
```

The variance of contribution amounts for transactions labeled "NA" is huge -
there's an 8,000-dollar difference between the biggest refund and the maximum
contribution of 2,700 dollars. More noteworthy than these outlying values is
the fact that the mean and median contribution amounts for contributions not
associated with an election type is considerably higher than the overall mean
and median contribution levels or the mean and median for either major-party
candidate. Over 400 of these "NA" contributions went to Trump's campaign.

Let's look at a quick scatterplot of campaign contributions over time colored
by the type of election. In order to keep the scatterplot from being
overplotted I'll set the transparency to 1/100 as I have with earlier
scatterplots.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contbs_by_Tp_Scatterplot}
# Create scatterplot of contributions over time colored by election type
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt,
           colour = election_tp), data = ga) +
  # Set transparency to 1/100
  geom_point(alpha = 0.01) +
  # Show contributions between 0 and 2700 without removing any data
  coord_cartesian(ylim = c(0, 2700)) +
  # Override the transparency aesthetic for key so colors are visible
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

As the dates progress from left to right, we see that the colors change from
turquoise, representing the primary election, to vermilion, representing the
general election. The variance is not visibly different between these two
elections, but the area between about 500 and 1,000 dollars becomes more filled
in toward the end of 2016 as Election Day approaches in early November.

It's not totally clear whether election type is strictly associated with date,
especially considering the "other" 2020 category that that one contribution is
labeled with. I'll look at a summary of contribution date by election type.

```{r echo=FALSE, message=FALSE, warning=FALSE, Receipt_Date_by_Election_Type}
by(ga$contb_receipt_dt, ga$election_tp, summary)
```

General and primary election contributions start showing up as early as October
2013, it turns out. Even though it's hard to see any real crossover in the
scatterplot there are some general election contributions sprinkled in during
the period before when the general election actually began.

I'd like to take a look at the data in the previous graph using a `geom_smooth`
with a linear model method. Maybe smooth conditional means will be helpful for
understanding contribution amounts over time as they relate to election type.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contbs_by_Election_Type_Smooth_LM}
# Plot smooth conditional means of contributions over time by election_tp
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt), data = ga) +
  geom_smooth(aes(colour = election_tp), method = 'lm')
```

The smoothers are a helpful but slightly oversimplified representation of the
contribution data over time. It's obvious that there is high variance in general
electioncontribution amounts in 2015, at which point many candidates are
starting to announce their intentions of running in 2016. As the data enters
2016, the means have much less variance, and the mean contribution to the
general election remains a bit higher on average than the mean contribution to
the primary election. The mean and median contribution amounts vary from
candidate to candidate but these smoothers give a sense of the gradual decrease
in mean contribution amounts as nomination solidify and more and more people
contribute to the various campaigns.

The approximately 450 contributions with election type "NA" have quite a bit of
variance, and seem to increase in mean amount as Election Day approaches.
Additionally, the handful of "other" 2016 contributions are represented by a
short, nearly vertical green line that reflects the high mean contribution
amount associated with those contributions.

I think what accounts for the high variance visible on the left side of the
graph of general election contributions is the relatively small number of
contributions coming in at that time. One or a few high contributions around
this time could set the mean and median as high as 2,700 dollars if that's the
only contribution received that day.

As more people start getting engaged with the campaigns underway, the mean comes
down. I think this may be because, as mentioned earlier, most people don't have
enough expendable income to pledge high amounts of money to presidential
campaigns, even if they want to.

## Candidate Party and Gender vs. Contribution Amount
After loading in the dataset I added columns for candidate party and gender.
I'd like to take a quick look at whether there are noticeable differences in
contribution amounts based on gender and political party.

The fact that Trump's mean and median contribution amounts both inside and
outside of Atlanta are so much higher than Clinton's leads me to guess that
contributions to the Republican candidates have higher amounts on average
than contributions to Democratic candidates.

First I'll use the `by` function to get an idea of what the distributions look
like.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contbs_by_Party_Summary}
# Summarize contribution amounts by candidate party
by(ga$contb_receipt_amt, ga$cand_party, summary)
```

I'll also look at a scatterplot with transparency added and zoomed in on
contributions under 500 dollars.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contbs_by_Party_Scatterplot}
# Create scatterplot of contributions colored by candidate party
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt,
           colour = cand_party), data = ga) +
  # Add transparency of 1/100
  geom_point(alpha = 0.01) +
  # Restrict x-axis to 2016 and y-axis to contributions under 500
  coord_cartesian(xlim = c(as.Date('2016-01-01'), as.Date('2016-12-31')),
                  ylim = c(0, 500)) +
  # Do not add transparency to plot's legend
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

Democrats have the lowest median and mean contribution amounts by far of all the
parties represented in the data set. The median amount (25 dollars) is half that
of the Republican median of 50 dollars, while the mean contribution amount for
Republican candidates is 72 dollars higher than the mean contribution to
Democrats.

As for third-party contributions, Green, Independent, and Libertarian candidates
all have a median contribution level of 100. The Libertarian candidate, Gary
Johnson, has the highest mean contribution amount at 224 dollars, followed 
closely by Jill Stein at 219.80. 

I'd also like to look at how contribution amounts break down by gender. Once
again I'll run the `by` function and plot a scatterplot with a transparency of
1/100, zooming in on contributions of 500 dollars or less.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contbs_by_Gender_Summary}
# Summarize contribution amounts by candidate gender
by(ga$contb_receipt_amt, ga$cand_gender, summary)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Contbs_by_Gender_Scatterplot}
# Create scatterplot of contributions colored by candidate gender
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt,
           colour = cand_gender), data = ga) +
  # Add transparency of 1/100
  geom_point(alpha = 0.01) +
  # Restrict x-axis to the year 2016 and y-axis to contributions under 500
  coord_cartesian(xlim = c(as.Date('2016-01-01'), as.Date('2016-12-31')),
                  ylim = c(0, 500)) +
  # Do not apply transparency to colors in legend 
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

The median contribution to female candidates is 25 dollars while the median
contribution to male candidates is 40 dollars. Mean contributions are higher
for men as well: 122.5 dollars to men versus 104.3 dollars to women. Once again
it's not immediately apparent whether the lower mean and median for the female
candidates has to do with Clinton's high number of contributions, or if it is
due rather to her party affiliation or some other cause.

## Atlanta vs. the Rest of Georgia
I'm curious how contributions from voters in Atlanta compared to contributions
from the rest of the state. There are 10 counties that are considered to be part
of the metro Atlanta area, but I'm only going to deal with contributions from
the City of Atlanta for the purposes of this investigation.

It'd be simplest to create subsets of the data by setting the city name equal to
or not equal to 'Atlanta'. However, in a quick glance at the data, it looks like
there are some misspellings of the city names here and there. So what I'll do is
create a subset of Atlanta contributions using the contributor zip
code column, `contbr_zip`.

I'll use the `dplyr` package to do some subsetting and reshaping of the data.
I'll also filter the data to contributions to Trump and Clinton, convert the zip
code column to numeric format, and truncate the entries in the zip code column to
five digits.
                  
The [USPS website](https://tools.usps.com/go/ZipLookupResultsAction!input.action?items=30&page=1&companyName=&address1=&address2=&city=Atlanta&state=GA&zip=) has a complete list of zip codes for the city of Atlanta. I'll
use this list to create a vector containing all of the Atlanta zip codes. I'll
also add a column containing "TRUE" or "FALSE" depending on whether the zip code
associated with the contribution was in the City of Atlanta or not, and filter
the dataframe to create two subsetted data frames containing contributions from
Atlanta and contributions from outside of Atlanta.

```{r echo=FALSE, message=FALSE, warning=FALSE, Zip_Code_Data}
# Use dplyr to create data.frame containing contributions to Clinton & Trump
ga_zips <- filter(ga, cand_nm == "Clinton, Hillary Rodham"
                  | cand_nm == "Trump, Donald J.")
# Truncate zip codes in dataset to first five digits
ga_zips$contbr_zip <- as.numeric(substr(ga_zips$contbr_zip, 0, 5))
# Reduce data.frame to transaction ID, candidate name, zip code, amount
ga_zips <- select(ga_zips, tran_id, cand_nm, contbr_zip, contb_receipt_amt)

# Character vector containing all of Atlanta's zip codes
atlanta_zip_codes <- c(30301, 30302, 30303, 30304, 30305, 30306, 30307, 30308,
                       30309, 30310, 30311, 30312, 30313, 30314, 30315, 30316,
                       30317, 30318, 30319, 30320, 30321, 30322, 30324, 30325,
                       30326, 30327, 30328, 30329, 30331, 30332, 30333, 30334,
                       30336, 30337, 30338, 30339, 30340, 30341, 30342, 30343,
                       30344, 30345, 30346, 30348, 30349, 30350, 30353, 30354,
                       30355, 30356, 30357, 30358, 30359, 30360, 30361, 30362,
                       30363, 30364, 30366, 30368, 30369, 30370, 30371, 30374,
                       30375, 30377, 30378, 30380, 30384, 30385, 30388, 30392,
                       30394, 30396, 30398, 31106, 31107, 31119, 31126, 31131,
                       31136, 31139, 31141, 31145, 31146, 31150, 31156, 31192,
                       31193, 31195, 31196, 39901)

# Add new column with "TRUE" if zip code is in Atlanta, "FALSE" if not
ga_zips <- mutate(ga_zips,
                  atlanta_zip = contbr_zip %in% atlanta_zip_codes)

# Create subsets of mutated data based on whether in Atlanta or not
atl_major_contributions <- filter(ga_zips, atlanta_zip == TRUE)
non_atl_major_contributions <- filter(ga_zips, atlanta_zip == FALSE)
```

Next I'll create a set of two dataframes for Atlanta contributions to Clinton
and Atlanta contributions to Trump, and run some statistical summaries on them.

```{r echo=FALSE, message=FALSE, warning=FALSE, Subset_ATL_Contributions}
# Create subsets of Atlanta contributions for each major-party candidate
atl_major_contributions.clinton <- subset(atl_major_contributions,
                                          cand_nm == "Clinton, Hillary Rodham")
atl_major_contributions.trump <- subset(atl_major_contributions,
                                        cand_nm == 'Trump, Donald J.')
```

Summary of contributions to Clinton from Atlanta:

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_Contribs_Clinton_Summary}
# Display summary of contributions to Clinton from Atlanta
summary(atl_major_contributions.clinton$contb_receipt_amt)
```

Count of contributions to Clinton from Atlanta:

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_Contribs_Clinton_Count}
# Display count of contributions to Clinton from Atlanta
dim(atl_major_contributions.clinton)
```

Sum of contributions to Clinton from Atlanta:

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_Contribs_Clinton_Sum}
# Display sum of contributions to Clinton from Atlanta
sum(atl_major_contributions.clinton$contb_receipt_amt)
```

Summary of contributions to Trump from Atlanta:

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_Contribs_Trump_Summary}
# Display summary of contributions to Trump from Atlanta
summary(atl_major_contributions.trump$contb_receipt_amt)
```

Count of contributions to Trump from Atlanta:

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_Contribs_Trump_Count}
# Display count of contributions to Trump from Atlanta
dim(atl_major_contributions.trump)
```

Sum of contributions to Trump from Atlanta:

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_Contribs_Trump_Sum}
# Display sum of contributions to Trump from Atlanta
sum(atl_major_contributions.trump$contb_receipt_amt)
```

Clinton received almost 24,000 contributions from Atlanta whereas Trump received
only 2,913 contributions from Atlanta. However, the median and mean contribution
amounts for Trump are considerably higher than those for Clinton:

* Trump received a median contribution of 100 dollars from Atlanta while Clinton
received a median contribution of 37 dollars.
* Trump received a mean contribution of 336.3 dollars from Atlanta while Clinton
receive a mean contribution of 158.6 dollars.
* Trump received a total of 979,781.7 dollars from Atlanta while Clinton
received a total of 3,802,396 dollars.

I'd like to look at histograms of the number of contributions from Atlanta to
the major-party candidates.

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_Contribs_Hist1}
# Create histograms of contributions from Atlanta to major-party candidates
qplot(x = contb_receipt_amt, 
      data = atl_major_contributions) +
  # Facet graph by candidate name and make it two columns
  facet_wrap(~cand_nm, ncol=2) +
  # Add title
  ggtitle("Atlanta Contributions")
```

It's a bit hard to see what's going on in this distribution so I'm going to
apply a log10 scale and see if that helps with understanding it.

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_Contribs_Hist2}
# Create histograms of contributions from Atlanta to major-party candidates and
# plot them on a log10 x-scale.
qplot(x = contb_receipt_amt, 
      data = atl_major_contributions) +
  scale_x_log10() +
  # Facet graph by candidate name and make it two columns
  facet_wrap(~cand_nm, ncol=2) +
  # Add title
  ggtitle("Atlanta Contributions")
```

Here we have two normal-ish distributions for Clinton and Trump. These
distributions reflect the fact that Clinton received way more individual
contributions than Trump did in Atlanta. It's a bit harder to tell that the mean
and median contribution amounts were greater for Trump but we can see that his
distribution is a bit further to the right on the x-axis than Clinton's is.

Next let's look at a boxplot of contributions from Atlanta to both major-party
candidates.

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_Contbs_Boxplots}
# Create boxplots of contributions from Atlanta to both major-party candidates
qplot(x = cand_nm, y = contb_receipt_amt,
      data = atl_major_contributions) +
    stat_summary(fun.y = mean, geom = 'point', colour = 'red', shape = 4) +
  geom_boxplot() +
  # Add title
  ggtitle("Atlanta Contributions")
```

Without zooming in, we can see that there is more variance in receipt amounts
for Trump than there is for Clinton, whose contributions remain in the -2,700-
to 2,700-dollar range. Let's zoom in using `coord_cartesian` to get a closer
look at the interquartile range without removing any values from the data. I'll
also add a red "x" using `stat_summary` marking the mean contribution amount for
each candidate.

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_Contbs_Boxplots_Zoom}
# Create boxplots of contributions from Atlanta to both major candidates
qplot(x = cand_nm, y = contb_receipt_amt,
      data = atl_major_contributions) +
  geom_boxplot() +
  # Add an 'X' marking the mean contribution amount for each candidate
  stat_summary(fun.y = mean, geom = 'point', colour = 'red', shape = 4) +
  # Zoom in to the area between 0 and 400 dollars without removing data
  coord_cartesian(ylim = c(0, 400)) +
  # Add title
  ggtitle("Atlanta Contributions")
```

This zoomed-in version makes it clear that most of Clinton's contributions from
Atlanta were between 25 dollars and 100 dollars each while most of the
contributions to Trump were much higher, between 50 and 250 dollars each. The
median and mean contribution amounts are visibly much higher for Trump than for
Clinton, but the thickness of the Clinton boxplot's whiskers reflects the
higher quantity of contributions she received.

Now let's take a look at contributions from outside Atlanta.

```{r echo=FALSE, message=FALSE, warning=FALSE, Non_ATL_Contributions_Code}
# Create subsets of non-Atlanta contributions for each major-party candidate
non_atl_major_contributions.clinton <- subset(non_atl_major_contributions,
                                          cand_nm == "Clinton, Hillary Rodham")
non_atl_major_contributions.trump <- subset(non_atl_major_contributions,
                                        cand_nm == 'Trump, Donald J.')
```

Summary of contributions to Clinton from Georgia (outside Atlanta):

```{r echo=FALSE, message=FALSE, warning=FALSE, Non_ATL_Contbs_Clinton_Summary}
# Display summary of contributions to Clinton from outside Atlanta
summary(non_atl_major_contributions.clinton$contb_receipt_amt)
```

Count of contributions to Clinton from Georgia (outside Atlanta):

```{r echo=FALSE, message=FALSE, warning=FALSE, Non_ATL_Contbs_Clinton_Count}
# Display count of contributions to Clinton from outside Atlanta
dim(non_atl_major_contributions.clinton)
```

Sum of contributions to Clinton from Georgia (outside Atlanta):

```{r echo=FALSE, message=FALSE, warning=FALSE, Non_ATL_Contbs_Clinton_Sum}
# Display sum of contributions to Clinton from outside Atlanta
sum(non_atl_major_contributions.clinton$contb_receipt_amt)
```

Summary of contributions to Trump from Georgia (outside Atlanta):

```{r echo=FALSE, message=FALSE, warning=FALSE, Non_ATL_Contbs_Trump_Summary}
# Display summary of contributions to Trump from outside Atlanta
summary(non_atl_major_contributions.trump$contb_receipt_amt)
```

Count of contributions to Trump from Georgia (outside Atlanta):

```{r echo=FALSE, message=FALSE, warning=FALSE, Non_ATL_Contbs_Trump_Count}
# Display count of contributions to Trump from outside Atlanta
dim(non_atl_major_contributions.trump)
```

Sum of contributions to Trump from Georgia (outside Atlanta):

```{r echo=FALSE, message=FALSE, warning=FALSE, Non_ATL_Contbs_Trump_Sum}
# Display sum of contributions to Trump from outside Atlanta
sum(non_atl_major_contributions.trump$contb_receipt_amt)
```

The contrasts between contribution amounts received have stark contrasts outside
of Atlanta. The mean and median contribution amounts for Trump are twice those
of Clinton. While Clinton received 44,419 contributions outside Atlanta, Trump
received 26,777 contributions and out-fundraised her by over 700,000 dollars.
I'll look at some visual representations of this.

```{r echo=FALSE, message=FALSE, warning=FALSE, NonATL_Contbs_Initial_Hist}
# Create histogram of contributions from outside Atlanta to major-party
# candidates, faceted by candidate name.
qplot(x = contb_receipt_amt, 
      data = non_atl_major_contributions) +
  facet_wrap(~cand_nm, ncol=2) +
  # Add title
  ggtitle("Non-Atlanta Contributions")
```

This first stab at a histogram of non-Atlanta contributions is a bit hard to
read so I'll do a log10 transformation of the x-axis scale as I did with the
distributions of the Atlanta contributions.

```{r echo=FALSE, message=FALSE, warning=FALSE, NonATL_Contbs_Histograms}
# Create histogram of contributions from outside Atlanta to major-party
# candidates, faceted by candidate name, on log10 x-scale.
qplot(x = contb_receipt_amt, 
      data = non_atl_major_contributions) +
  scale_x_log10() +
  facet_wrap(~cand_nm, ncol=2) +
  # Add title
  ggtitle("Non-Atlanta Contributions")
```

Compared to the distribution of his Atlanta contributions, Trump has a more
robust normal distribution of individual contributions received from outside of
Atlanta. The y-scale of the count of contributions received is much higher as
well for both distributions. It'd be helpful to look at these contributions all
on the same scale, so I'll save the plots I previously made into variables and
use the `grid.arrange` function from the `gridExtra` package to look at them all
at once.

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_v_NonATL_Contbs_Hist}
# Save previous histogram of major-party contributions from Atlanta.
# Put graph on y-scale of 0 to 10,000 for comparison with non-ATL contributions.
# Add title to plots.
atl_maj_con_compare <- qplot(x = contb_receipt_amt, 
      data = atl_major_contributions) +
  scale_x_log10() +
  ylim(0, 10000) +
  facet_wrap(~cand_nm, ncol=2) +
  ggtitle('Atlanta Contributions')

# Save previous histogram of major-party contributions from outside Atlanta.
# Put graph on y-scale of 0 to 10,000 for comparison with ATL contributions.
# Add title to plots.
non_atl_maj_con_compare <- qplot(x = contb_receipt_amt, 
      data = non_atl_major_contributions) +
  scale_x_log10() +
  ylim(0, 10000) +
  facet_wrap(~cand_nm, ncol=2) +
  ggtitle('Non-Atlanta Contributions')

# Show all the histograms at once
grid.arrange(atl_maj_con_compare, non_atl_maj_con_compare, ncol = 1)
```

Looking at everything at once, it's apparent that Trump did way better outside
of Atlanta than he did inside Atlanta. This seems to fit with what I've heard
about Atlanta skewing much more liberal than conservative compared to the rest
of the state. 

Let's also look at boxplots of the non-Atlanta contributions for comparison.

```{r echo=FALSE, message=FALSE, warning=FALSE, Non_ATL_Contbs_Boxplots}
# Create boxplots of contributions to major-party candidates from
# outside of Atlanta.
qplot(x = cand_nm, y = contb_receipt_amt,
      data = non_atl_major_contributions) +
  geom_boxplot() +
  # Add title
  ggtitle("Non-Atlanta Contributions")
```

This initial boxplot reflects the high variance of contributions to Trump from
outside of Atlanta, with amounts ranging from -5,900 to 12,500. Let's zoom in
on the interquartile ranges and get a better look at where most of the
distribution's values lie.

```{r echo=FALSE, message=FALSE, warning=FALSE, Non_ATL_Contbs_Boxplots_Zoom}
# Create boxplots of contribution amounts to major-party candidates from
# outside of Atlanta.
qplot(x = cand_nm, y = contb_receipt_amt,
      data = non_atl_major_contributions) +
  geom_boxplot() +
  # Add an 'X' marking the mean contribution amount for each candidate
  stat_summary(fun.y = mean, geom = 'point', colour = 'red', shape = 4) +
  # Zoom into the area between 0 and 400 without removing data
  coord_cartesian(ylim = c(0, 400)) +
  # Add title
  ggtitle("Non-Atlanta Contributions")
```

As with contributions from Atlanta, the mean and median contribution amounts for
Trump are much higher than those of Clinton when it comes to contributions in
the rest of the state. There is also considerably more variance in contribution
amounts for Trump than for Clinton.

I'd like to look at all the boxplots at once and see if that provides any
further insight.

```{r echo=FALSE, message=FALSE, warning=FALSE, ATL_Contribs_Compare}
# Save boxplots of Atlanta major-party contributions to variable
atl_maj_con_boxplot <- qplot(x = cand_nm, y = contb_receipt_amt,
      data = atl_major_contributions) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', colour = 'red', shape = 4) +
  coord_cartesian(ylim = c(0, 400)) +
  ggtitle("Atlanta Contributions")

# Save boxplots of non-Atlanta major-party contributions to variable
non_atl_maj_con_boxplot <- qplot(x = cand_nm, y = contb_receipt_amt,
      data = non_atl_major_contributions) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', colour = 'red', shape = 4) +
  coord_cartesian(ylim = c(0, 400)) +
  ggtitle("Non-Atlanta Contributions")

# Display boxplots one above the other
grid.arrange(atl_maj_con_boxplot, non_atl_maj_con_boxplot, ncol = 1)
```

It's notable that the shapes of the boxplots have similar size ratios for both
Atlanta and non-Atlanta contributions. The mean and median contributions to
both candidates are higher in Atlanta than they are statewide. I'm not sure if
this means that Atlanta residents are more affluent on average than people
outside the city, or if it's the larger number of contributions from outside
Atlanta (approximately 70,000 versus the approximately 26,000 from Atlanta) that
brings down the mean and median.

### Atlanta/Non-Atlanta Form Types
I'm curious whether there are any associations between the form type and
location in terms of Atlanta vs. non-Atlanta contributions.

```{r echo=FALSE, message=FALSE, warning=FALSE, Form_Type_vs_Location}
# Create smaller data.frame from overall data containing zip code, transaction
# ID, form type, and party. Truncate zip code to five digits.
ga_form_data <- select(ga, contbr_zip, tran_id, form_tp, cand_party)
ga_form_data$contbr_zip <- substr(ga_form_data$contbr_zip, 0, 5)
# Add a column to this data.frame which says "TRUE" if contribution is in
# Atlanta and false if contribution is from outside Atlanta.
ga_form_data <- mutate(ga_form_data,
                       atl = contbr_zip %in% atlanta_zip_codes)

# Look at summary of form types by whether address is in Atlanta or not
by(ga_form_data$form_tp, ga_form_data$atl, summary)
```

There were almost three times the number of campaign contributors from
outside Atlanta than from inside Atlanta. Additionally there were more
than four times as many "authorized committee" contributions from outside of
Atlanta (29,776) than there were from Atlanta (6,878). 

## Form Type vs. Contribution Amount Over Time
I'm wondering if there is any connection between form type and contribution
amount.

As I saw in the univariate analysis, most of the contributions (116,284 out of
153,985) are classed as "Form 3P Line 17A", meaning they are contributions from
individuals, not committees. There are, however, 36,654 contributions designated
as "Form 3P Line 17A" (transfers from authorized committees) and 1,047
transactions categorized as "Form 3P Line 28A (refunds to individuals).

I'll look at some quick summaries and scatterplots to see if that tells me
anything about a possible relationship between form type and contribution
amount.

```{r echo=FALSE, message=FALSE, warning=FALSE, Form_Tp_vs_Contb_Amt}
# Summarize contribution amounts by form type
by(ga$contb_receipt_amt, ga$form_tp, summary)
```

Contributions categorized as being from authorized committees have higher mean
and median contribution amounts - a mean of 152.5 dollars and a median of 50
dollars - than contributions categorized as being from individuals, which have
a mean of 106 dollars and a median of 27 dollars.

I'll look at contributions over time colored by form type. I'll initially focus
in on the area between 0 and 2700 dollars for the year 2016, and set the
transparency low so I can get a better sense of what's going on here.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contbs_Time_Form_Type_SPlot_Zoom}
# Create scatterplot of contributions over time by contribution amount
# Color points by form type
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt, colour = form_tp),
       data = ga) +
  # Set transparency to 1/100
  geom_point(alpha = 0.01) +
  # Limit x-axis to year 2016 and y-axis to contributions under 2700 dollars
  coord_cartesian(xlim = c(as.Date('2016-01-31'), as.Date('2016-12-31')),
                  ylim = c(0, 2700)) +
  # Keep the colors fully visible in the legend
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

It looks like the variance is mostly comparable for form types 17A and 18, and
it's no surprise that all the form 28A transactions are negative values as they
are refunds.

Committee contributions are visibly a bit higher on average than individual
"SA17A" contributions. Interestingly the committee contributions don't really
gear up in earnest until July 2016 when the party conventions are going on.

The overall variance in contribution amounts for committee contributions is not
that much greater than that for regular individual contributions. Additionally
the vast majority of the contributions appear to be less than 100 dollars.

I'll zoom in just a bit closer to the area between 0 and 250 dollars and see if
that adds anything to my understanding.

```{r echo=FALSE, message=FALSE, warning=FALSE, Contbs_Time_Form_Type_SPlot_Zoom_100}
# Create scatterplot of contributions over time by contribution amount
# Color points by form type
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt, colour = form_tp),
       data = ga) +
  # Set transparency to 1/100
  geom_point(alpha = 0.01) +
  # Limit x-axis to year 2016 and y-axis to contributions under 250 dollars
  coord_cartesian(xlim = c(as.Date('2016-01-31'), as.Date('2016-12-31')),
                  ylim = c(0, 250)) +
  # Keep the colors fully visible in the legend
  guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

I can see here that most committee contributions tend to be at least 25 dollars
and that most of the green points representing these contributions have, as I
saw before, a higher average contribution amount.

Are there any details about the committees themselves to be found in the data?
I'll take a quick look at the first rows of filtered data.frame showing
only committee contributions.

```{r echo=FALSE, message=FALSE, warning=FALSE, Committee_Contributions}
# Filter the ga data.frame to show only rows with form type 18A
sa18 <- filter(ga, form_tp == 'SA18')
# Look at the first ten rows of this new data.frame
head(sa18, 10)
```

Other than the form type these contributions don't look terribly different from
individual contributions - there's an individual name, city, state, zip code,
occupation, and employer associated with each of these contributions.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
I observed once again that mean and median contributions to Trump and other
Republican candidates were much higher than mean and median contributions
to Democratic candidates. It's also clear that, while Clinton had lower
contribution amounts on average, she received way more individual
contributions throughout the state than Trump did.

There also seemed to be relationships between location and both contribution
amounts and candidate contributed to. Atlanta contributions are higher on 
average than contributions from outside of Atlanta. Additionally contributions
from Atlanta appear much more likely to contribute to Clinton instead of Trump.
While Clinton enjoyed a considerable amount of contributions from the rest of 
the state, it's the rest of the state where Trump did most of his fundraising.
Additionally, it appears that there is a relationship between form type and
party as well as form type and location (i.e. in or outside of Atlanta).

Looking at daily contribution dollar amount totals turned out to be very useful.
I created line graphs of daily contribution totals broken down by candidate
party and candidate gender. These graphs, when overlaid with vertical lines
reflecting dates of debates and of a few significant campaign events, provided
different insights than when I looked at these events in relation to a
scatterplot of contributions over time.

Contributions from authorized committees tended to have higher dollar amounts
than contributions from individuals. Additionally, it looked like contributions
from committees appeared to start coming in more rapidly once the general
election was underway.

It did not appear that contribution amounts varied a great deal with election
type. Some of the earliest contributions were very high but for the most part
individual contributions were less than 100 dollars apiece.

### Were there any interesting or surprising interactions between features?
When I looked at daily contribution totals by party and added vertical lines
representing debates and turning points in the election, all of a sudden
I was able to get a better sense of what the effects of those events may
have been. This was not immediately apparent in the earlier parts of my
explorations. Additionally, what these graphs with the events overlaid on them
suggested is that the events had a different effect than I would have expected.

The difference between mean and median contribution amounts to primary and
general elections was not immediately visible on a scatterplot or in summaries
of the overall data. However, it turned out that Trump received much higher
contribution amounts on average during the general election than he did during
the primary election. Clinton's average contribution amount went
down once the general election began. The earlier the date of an election
contribution the more variance there was observable in median contribution
amount.

Contributions to male candidates were significantly higher in mean and median
contribution amounts than were contributions to female candidates. However there
were points during the primary election where female candidates had higher daily
totals than did the male candidates, which is interesting in light of how many
male candidates were still in the running at that point.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
# Create line graphs of contribution counts by day to Clinton & Trump
# Save plot to a variable
contb_counts_final <- qplot(x = contb_receipt_dt, 
      data = subset(ga, cand_nm == 'Clinton, Hillary Rodham' | 
                      cand_nm == 'Trump, Donald J.'),
      binwidth = 1, geom = 'freqpoly', colour = cand_nm) +
  # Restrict x-axis to June-December 2016, minor breaks by month
  scale_x_date(limits = c(as.Date('2016-06-01'), as.Date('2016-12-31')),
               date_minor_breaks = "1 month") +
  # Set x-axis title
  xlab("Contribution Receipt Date") +
  # Set y-axis title
  ylab("Number of Contributions") +
  # Set title for plot
  ggtitle(paste("Contributions from Georgia to Major-Party Nominees",
          "U.S. Presidential Election, 2016", sep = "\n"))

# Display the graph
contb_counts_final +
  # Use Paul Tol color scheme from ggthemes package
  scale_color_ptol("cyl") +
  # Add minimal theme
  theme_minimal() +
  # Remove the legend title
  theme(legend.title = element_blank())
```

### Description One
This plot shows how the number of individual contributions from Georgia to the
two major-party candidates changed over the course of the election, after the
primaries. During June, July, and August, there are several spikes in
contributions to Trump. This is the time period in which the pool of Republican
nominees shook down and the Republican National Convention took place. However,
Clinton received a higher number of contributions day-to-day from August through
Election Day. Once Election Day arrives, contributions to Clinton end but Trump
receives a trickle of contributions through early December.

The reason I chose this plot is because it surprised me that Clinton received
more individual contributions day-to-day for much of general election season. It
also gives an idea of how volatile the day-to-day numbers of contributions are.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Contbs_Scatter_Jun_Dec_Final}
# Create plot showing all contributions to Trump during general election
# Save to variable
trump_contbs_time_general_final <- ggplot(aes(x = contb_receipt_dt,
                                        y = contb_receipt_amt),
                                    data = trump) +
  # Jitter scatterplot points and color them red
  geom_jitter(alpha = 0.01, colour = "red") +
  # Set x-axis limits to 6/15/16-12/15/16 with breaks at start of each month
  scale_x_date(limits = c(as.Date('2016-06-15'), as.Date('2016-12-15')),
               breaks = seq(as.Date('2016-06-01'), as.Date('2016-12-31'),
                            'month')) +
  # Set limit on y-axis of 0 to 500 dollars
  ylim(0, 500) +
  # Label the y-axis
  ylab("Amount ($)") +
  # Title the plot
  ggtitle("Donald Trump") +
  # Make title text size 12
  theme(plot.title = element_text(size = 12),
        # Make axis title text size 10
        axis.title = element_text(size = 10),
        # Remove the x-axis title
        axis.title.x = element_blank())

# Create plot showing all contributions to Clinton during general election
# Save to variable
clinton_contbs_time_general_final <- ggplot(aes(x = contb_receipt_dt,
                                          y = contb_receipt_amt),
                                  data = clinton) +
  # Jitter the scatterplot points and color them blue
  geom_jitter(alpha = 0.01, colour = "blue") +
  # Set x-axis limits to 6/15/16-12/15/16 with breaks at start of each month
  scale_x_date(limits = c(as.Date('2016-06-15'), as.Date('2016-12-15')),
               breaks = seq(as.Date('2016-06-01'), as.Date('2016-12-31'),
                            'month')) +
  # Set limit on y-axis of 0 to 500 dollars
  ylim(0, 500) +
  # Label the y-axis
  ylab("Amount ($)") +
  # Title the plot
  ggtitle("Hillary Clinton") +
  # Make title text size 12
  theme(plot.title = element_text(size = 12),
        # Make axis title text size 10
        axis.title = element_text(size = 10),
        # Remove the x-axis title
        axis.title.x = element_blank())

# Create scatterplot showing all contributions during general election period
# Save to variable
all_contbs_time_general_final <- ggplot(aes(x = contb_receipt_dt,
                                      y = contb_receipt_amt), data = ga) +
  # Jitter scatterplot points and color them purple
  geom_jitter(alpha = 0.01, colour = "purple") +
  # Set x-axis limits to 6/15/16-12/15/16 with breaks at start of each month
  scale_x_date(limits = c(as.Date('2016-06-15'), as.Date('2016-12-15')),
               breaks = seq(as.Date('2016-06-01'), as.Date('2016-12-31'),
                            'month')) +
  # Set limit on y-axis of 0 to 500 dollars
  ylim(0, 500) +
  # Label the y-axis
  ylab("Amount ($)") +
  # Title the plot
  ggtitle("All Candidates") +
  # Make title text size 12
  theme(plot.title = element_text(size = 12),
        # Make axis title text size 10
        axis.title = element_text(size = 10),
        # Remove the x-axis title
        axis.title.x = element_blank())

# Plot all three scatterplots together in one column
grid.arrange(trump_contbs_time_general_final,
  clinton_contbs_time_general_final,
  all_contbs_time_general_final, ncol = 1,
  # Set title for all three plots
  top = paste("Individual Contributions Under $500 from Georgia",
              "to U.S. Presidential Election, 2016 (General)", sep = "\n"))
```

### Description Two
Most contributions from Georgia to the various presidential candidates in the
2016 election were under 500 dollars. Looking at a scatterplot of these
contributions over time, with a transparency of 1/100, gives a sense of the ebb
and flow of contributions from Georgia residents during the general election
period.

Clinton visibly has a more sustained number of contributions coming in,
particularly at 100 dollars or less per contribution. Trump, on the other hand,
has hot and cold periods, with a less sustained intake but a visibly higher
average contribution amount, as well as more contributions at the 250- and
400-dollar levels.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Daily_Totals_by_Cand_Nm}
# Select columns from dataset: candidate name, date, amount, transaction ID
daily_nominee_totals_cols <- select(ga, cand_nm, contb_receipt_dt, 
                                  contb_receipt_amt, tran_id)

# Filter data.frame to Democratic, Republican, Green, and Libertarian nominees
daily_nominee_totals_cols <- filter(daily_nominee_totals_cols,
                                    cand_nm == "Trump, Donald J."
                                    | cand_nm == "Clinton, Hillary Rodham"
                                    | cand_nm == "Stein, Jill"
                                    | cand_nm == "Johnson, Gary")

# Group columns in new data.frame by name and receipt date
# Add new column for sum daily total contributions received by candidate
daily_nominee_totals <- daily_nominee_totals_cols %>% 
  group_by(cand_nm, contb_receipt_dt) %>% 
  summarise(daily_nominee_total = sum(contb_receipt_amt))

# Plot line graphs representing daily contribution totals to candidates by name
# Save this figure to a variable
daily_nominee_totals <- ggplot(aes(x = contb_receipt_dt,
                                   y = daily_nominee_total, colour = cand_nm),
                                   data = daily_nominee_totals) +
  geom_line() +
  # Add blue dashed line representing September 26 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-09-26'))),
             colour = 'blue', linetype = 4) +
  # Add purple dashed line representing October 9 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-09'))), 
             colour = 'purple', linetype = 4) +
  # Add red dashed line representing October 19 presidential debate
  geom_vline(aes(xintercept = as.numeric(as.Date('2016-10-19'))), 
             colour = 'red', linetype = 4) +
  # Restrict x-axis to September 1 through November 15
  scale_x_date(lim = c(as.Date('2016-09-01'), as.Date('2016-11-15')),
               date_minor_breaks = "1 month") +
  # Restrict y-axis to contributions between 0 to 90000
  scale_y_continuous(lim = c(0, 90000)) +
  # Specify title for graph
  ggtitle(paste("Total Daily Campaign Contributions from Georgia vis--vis",
                "U.S. Presidential Debates (9/26, 10/9, 10/19), 2016",
                sep = "\n")) +
  # Leave x-axis label blank
  xlab("") +
  # Specify y-axis label
  ylab("Daily Contribution Total ($)")

# Display the graph
daily_nominee_totals +
  # Use Paul Tol color scheme from ggthemes package
  scale_color_ptol("cyl") +
  # Add minimal theme
  theme_minimal() +
  # Set plot title text to size 11 and center it
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        # Remove legend title
        legend.title = element_blank(),
        # Set axis title text to size 10
        axis.title = element_text(size = 10))
```

### Description Three
This plot, which happens to look at potential election turning points, was
itself a turning point in my analysis process. I had been looking at individual
contribution amounts in relation to time, and struggled to find any kind of
connection between certain election events and the contribution amounts that
were coming in. Calculating the daily contribution totals and plotting them as
line graphs broken up by party gave a better sense of 

After each presidential debate there is all sorts of dialogue in the media
about who "won" each time. This is a difficult thing to quantify or assess with
any kind of objectivity, but looking at short-term impacts on daily campaign
fundraising at the state level might help give a regional picture of who "won."

In this case, we can observe a couple of things. There's a drop in contributions
from Georgia to Clinton right after the first debate while the fundraising
momentum that Trump had in Georgia prior to that debate continues to climb for
at least a few days afterward. A similar pattern emerges with the second debate:
Clinton's daily contribution totals drop while Trump's climb at that time
continues for a few days afterward. The third debate marks the beginning of yet
another drop in contributions from Georgia to Clinton, and while Trump's
contributions have a positive slope approaching that day, they decline sharply
afterward, hitting nearly the same low that Clinton hits.

It's also interesting to observe the small spikes in contributions to Gary
Johnson, the Libertarian nominee, and Jill Stein, the Green nominee. These
spikes don't appear to have too much of a connection to the debates, but there
is a small spike in Johnson's contributions after the first debate. Perhaps some
right-leaning viewers of the first debate didn't connect with Trump and decided
to contribute to the Libertarian candidate.

------

# Reflection
It turned out there were far more ways to break down state-level election
contributions than I thought. Even after looking at this data from all the
angles that I have, I feel like I've only really scratched the surface of what
took place over the course of the election from a campaign finance perspective.

The data posed some difficulties that I wasn't sure how to get around. One I'd
be particularly keen to resolve is consolidating contributions to the individual
level with the use of `dplyr` or another reshaping package. This seems almost
impossible to do given that contributor names and city names can be misspelled.
Additionally it's entirely possible to give multiple contributions and provide
correct demographic information as of a particular date that differs from
demographic information provided on another date - a contributor may have moved
cities or zip codes, started a new job or changed their last name after a
marriage, or started a new job.

It's also difficult to address the outliers in the contribution amounts column.
What makes this difficult are the reassignments that take place for compliance
with campaign finance regulations. There are also a wide variety of memos
associated with these contributions, the significance of which I did not have
the bandwidth to fully delve into.

As far as what went well, I found that I was able to create an interesting and
aesthetically appealing (to me, anyway) range of visuals that showed many
different ways of considering this data. Histograms, boxplots, frequency
polygons, line graphs, and scatterplots all offered windows to this data that
opened my eyes to what took place in this election.

I was surprised by a few things. Firstly I got the sense that both Clinton and
Sanders did far better from a fundraising standpoint than I thought they would
in light of the state's recent voting history. Secondly the effects of the
turning points I looked at - the presidential debates, the Billy Bush tape,
and Comey's letter to Congress - on daily contribution totals were not what I
would have expected.

There are future analyses I'd like to do with this dataset down the road. I'd
like to analyze contributors from Atlanta suburbs versus the city itself, and
also look at these two areas compared to the rest of the state. It'd also be
interesting to look at a wider range of turning points and attempt to quantify
the effect that these turning points did or did not have on the contribution
amounts and totals.